### 2024-07-23 manuscript.qmd
Added Bayes factor calculations

---
title: "Neural correlates of the deployment of spatial attention, and their modulation by repetitive movements"
author: "Cameron Smith & Daniel H. Baker"
bibliography: references.bib
csl: elife.csl
execute: 
  echo: false
  include: false
  output: false
format: pdf
---

```{r setup}
#| warning: false

processdata <- 0
# the processdata flag has three levels:
# 0 - just produce the manuscript with minimal analysis, using existing figures
# 1 - do group-level analysis using pre-computed MVPA and ERPs (requires 200MB)
# 2 - do all analyses from scratch using the raw data (requires ~35GB)

# install R packages
packagelist <- c('knitr','reticulate','osfr','bookdown','grImport','tiff','rstatix','ez','readr','BayesFactor','bmp','pracma')

missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

# list of participants who report stimming (for Expt 2)
stimlist <- c(201,204,210,211,212,214,216,217,218,219,220,221,223,224)
# list of participants who report not stimming (for Expt 2)
nostimlist <- c(205,206,207,208,209,215,222,225)

# create python environment and install packages
if (processdata==2){
if (!dir.exists('env')){
  system('python3 -m venv env')
  system('source env/bin/activate')
  system('python3 -m pip install mne numpy scikit-learn matplotlib pandas')
}
use_virtualenv('env/')
}

if (!dir.exists('local/processedE1')){dir.create('local/processedE1')}
if (!dir.exists('local/processedE2')){dir.create('local/processedE2')}

getcleanmean <- function(input){
  s <- dim(input)
  allsds <- apply(input,c(1,3),sd)
  thresh <- quantile(allsds,0.95)
  allmeans <- matrix(0,nrow=s[2],ncol=s[3])
  for (ch in 1:s[3]){
    temp <- allsds[,ch]
    i <- which(temp<thresh)
    allmeans[,ch] <- colMeans(input[i,,ch])
  }
return(allmeans)}

getcleanCI <- function(input){
  s <- dim(input)
  allsds <- apply(input,c(1,3),sd)
  thresh <- quantile(allsds,0.95)
  allCI <- matrix(0,nrow=s[2],ncol=s[3])
  for (ch in 1:s[3]){
    temp <- allsds[,ch]
    i <- which(temp<thresh)
    allCI[,ch] <- 1.96*apply(input[i,,ch],2,sd)/sqrt(length(i))
  }
return(allCI)}

flatalpha <- function(col, alpha=1){apply(sapply(col, col2rgb)/255, 2, function(x) rgb(x[1]*alpha + 1-alpha, x[2]*alpha + 1-alpha, x[3]*alpha + 1-alpha))}

addalpha <- function(col, alpha=1){apply(sapply(col, col2rgb)/255, 2, function(x) rgb(x[1], x[2], x[3], alpha=alpha))}

v4Interp <- function(df, xo, yo, rmax = .75, gridRes = 67) {
  ## Create a function to perform Matlab's v4 interpolation.
  ## Takes as input a data-frame with columns x, y, and z (x co-ordinates, y co-ordinates, and amplitude)
  ## and variables xo and yo, the co-ordinates which will be use to create a grid for interpolation
  xo <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
  yo <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
  xy <- df$x + df$y*sqrt(as.complex(-1))
  d <- matrix(rep(xy,length(xy)),nrow = length(xy), ncol = length(xy))
  d <- abs(d - t(d))
  diag(d) <- 1
  g <- (d^2) * (log(d)-1)   # Green's function.
  diag(g) <- 0
  weights <- qr.solve(g,df$z)
  xy <- t(xy)
  outmat <- matrix(nrow = gridRes,ncol = gridRes)
  for (i in 1:gridRes){
    for (j in 1:gridRes) {
      test4 <- abs((xo[i,j] + sqrt(as.complex(-1))*yo[i,j]) - xy)
      g <- (test4^2) * (log(test4)-1)
      outmat[i,j] <- g %*% weights}}
  outDf <- data.frame(x = xo[,1],outmat)
  names(outDf)[1:length(yo[1,])+1] <- yo[1,]
  return(outDf)}

mkgrating <- function(regionsize, f, o, p, c){
# ported from the Matlab function, originally by Tim Meese
# generates a single component sine wave grating using the following inputs:
  # regionsize is the width of the stimulus
  # f is spatial frequency in cycles per image
  # o is orientation in degrees
  # p is phase in degrees relative to the centre of the image
  # c is contrast

p <- p*(pi/180)   # convert phase from degrees to radians
o <- o*(pi/180)   # convert orientation from degrees to radians
f <- f/regionsize # scale frequency by image size
x0 <- (regionsize+1)/2  # locate centre of image
y0 <- x0

u <- f * cos(o) * 2*pi
v <- f * sin(o) * 2*pi

gridout <- meshgrid(1:regionsize,1:regionsize)   # generate x and y coordinate systems, requires the pracma package
xx <- gridout$X
yy <- gridout$Y

output <- matrix(0,nrow=regionsize,ncol=regionsize)
output <- (c * sin(u * (xx-x0) + v*(yy-y0) + p))

return(output)}

make_soft_window <- function(W,H,D=0.9){
  # Mark's function to make sine-wave gratings, ported from Matlab
  # W is width (in pixels), H is height (in pixels)
  # D is the diameter of the soft window at half height as a proportion of the width

  radius <- min(W*D/2,H*D/2)     # radius in pixels
  L <- 2*(min(W/2,H/2) - radius) # blur half-cycle
  X1 <- seq(-L/2,L/2)

  X <- (1:W) - (W/2)
  Y <- (1:H) - (H/2)
  xx <- matrix(rep(X,H),nrow=H,ncol=W,byrow=TRUE)
  yy <- t(matrix(rep(Y,W),nrow=W,ncol=H,byrow=TRUE))

  mask <- (xx^2 + yy^2)
  mask[mask<=(radius^2)] <- 1
  mask[mask>(radius^2)] <- 0

  WinKernel <- cos(X1*pi/L)  # half-cycle cosine
  convkernW <- (1:W)*0
  convkernW[(1+(W/2)-length(X1)/2):((W/2)+length(X1)/2)] <- WinKernel
  convkernH <- (1:H)*0
  convkernH[(1+(H/2)-length(X1)/2):((H/2)+length(X1)/2)] <- WinKernel
  cH <- t(apply(mask,2,convolve,convkernH))
  cH <- cH[,c((1+H/2):H,1:(H/2))]
  cW <- apply(mask,1,convolve,convkernW)
  cW <- cW[c((1+W/2):W,1:(W/2)),]
  mask <- cH * cW

  mask <- mask/max(mask)

  return(mask)}

# check for local files and directories and create/download if missing
if (!dir.exists('local')){dir.create('local')}
figdir <- 'Figures/'
if (!dir.exists(figdir)){dir.create(figdir)}

if (processdata>0){
if (!file.exists('local/MNEchannels.csv')){
osfproject <- osf_retrieve_node("dns89")
osffiles <- osf_ls_files(osfproject,n_max=300)
fid <- pmatch('MNEchannels.csv',osffiles$name)
osf_download(osffiles[fid,],path='local/')
}
if (!file.exists('local/channeldata.csv')){
osfproject <- osf_retrieve_node("dns89")
osffiles <- osf_ls_files(osfproject,n_max=300)
fid <- pmatch('channeldata.csv',osffiles$name)
osf_download(osffiles[fid,],path='local/')
}

chandata <- read.csv('local/channeldata.csv')
mnechans <- read.csv('local/MNEchannels.csv')
mnechans <- mnechans[c(1:12,14:18,20:64),1]  # exclude the mastoids
electrodeindices <- match(toupper(mnechans),chandata$Electrode)
}

if (processdata<2){
  # download processed data if missing
  if (!file.exists('local/Expt1resps.RData')){
osfproject <- osf_retrieve_node("dns89")
osffiles <- osf_ls_files(osfproject,n_max=300)
fid <- pmatch('Expt1resps.RData',osffiles$name)
osf_download(osffiles[fid,],path='local/')}
  if (!file.exists('local/Expt2resps.RData')){
osfproject <- osf_retrieve_node("dns89")
osffiles <- osf_ls_files(osfproject,n_max=300)
fid <- pmatch('Expt2resps.RData',osffiles$name)
osf_download(osffiles[fid,],path='local/')}
  if (!file.exists('local/Exp1Group.RData')){
osfproject <- osf_retrieve_node("dns89")
osffiles <- osf_ls_files(osfproject,n_max=300)
fid <- pmatch('Exp1Group.RData',osffiles$name)
osf_download(osffiles[fid,],path='local/')}
    if (!file.exists('local/Exp2Group.RData')){
osfproject <- osf_retrieve_node("dns89")
osffiles <- osf_ls_files(osfproject,n_max=300)
fid <- pmatch('Exp2Group.RData',osffiles$name)
osf_download(osffiles[fid,],path='local/')}
}

if (processdata==2){
  # check for presence of raw data and download if missing
  if (!dir.exists('local/rawdata')){dir.create('local/rawdata')}
  d <- dir('local/rawdata/',pattern='S')
  if (length(d)<126){
    osfproject <- osf_retrieve_node("yhsmb")
    osffiles <- osf_ls_files(osfproject,n_max=300)
    for (n in 1:nrow(osffiles)){
      if (!pmatch(osffiles[n,]$name,d,nomatch=0)){
        osf_download(osffiles[n,],path='local/rawdata/',progress=TRUE)
      }
    }
    osfproject <- osf_retrieve_node("dhkcg")
    osffiles <- osf_ls_files(osfproject,n_max=300)
    for (n in 1:nrow(osffiles)){
      if (!pmatch(osffiles[n,]$name,d,nomatch=0)){
        osf_download(osffiles[n,],path='local/rawdata/',progress=TRUE)
      }
    }    
  }
}

```

```{python ProcessExpt1}

# code to import raw EEG data and process

if r.processdata==2:

  import numpy as np
  import matplotlib.pyplot as plt
  import pandas as pd
  import mne
  from mne.decoding import (
      SlidingEstimator,
      GeneralizingEstimator,
      Scaler,
      cross_val_multiscore,
      LinearModel,
      get_coef,
      Vectorizer,
      CSP
  )

  from sklearn.pipeline import make_pipeline
  from sklearn.preprocessing import StandardScaler
  from sklearn.linear_model import LogisticRegression

  ANT_montage = mne.channels.make_standard_montage("standard_1020")
  reject_criteria = None   #dict(eeg=150e-6)

  clf = make_pipeline(StandardScaler(), LogisticRegression(solver="liblinear"))
  time_decod = SlidingEstimator(clf, n_jobs=10, scoring="roc_auc", verbose=True)

  legaltriggers = {
    "1": 1,
    "11": 2,
    "12": 3,
    "201": 4,
    "202": 5,
    "21": 6,
    "22": 7,
    "31": 8,
    "32": 9   
    }
  sublist = list(range(501,506)) + list(range(508,542))

  for subj in range(0,len(sublist)):
    filename = "local/rawdata/S" + str(sublist[subj]) + ".set"
    raw = mne.io.read_raw_eeglab(filename,preload=True)
    raw.drop_channels(["HEOG", "VEOG", "M1", "M2"])
    raw.set_montage(ANT_montage)

    filtered = raw.filter(0.2,30)

    ica = mne.preprocessing.ICA(n_components=20, random_state=97, max_iter=800)
    ica.fit(filtered)
    ica.exclude = [1, 2]  # details on how we picked these are omitted here
    ica.apply(filtered)

    events, _ = mne.events_from_annotations(filtered, event_id=legaltriggers)

    event_dict = {
        "cueleft": 4,
        "cueright": 5,
        "targetleft/top": 2,
        "targetleft/bottom": 3,
        "targetright/top": 6,
        "targetright/bottom": 7,
        "responsetop": 8,
        "responsebottom": 9
    }

    epochs = mne.Epochs(
        filtered,
        events,
        event_id=event_dict,
        tmin=-0.2,
        tmax=1,
        reject=reject_criteria,
        preload=True,
    )

    epochs.equalize_event_counts(["cueleft", "cueright"])
    cueepochs = epochs["cueleft", "cueright"]
    X = cueepochs.get_data()  # EEG signals: n_epochs, n_meg_channels, n_times
    y = cueepochs.events[:, 2]  # target: auditory left vs visual left

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    cuedecode = np.mean(scores, axis=0)

    epochs.equalize_event_counts(["targetleft", "targetright"])
    targetepochs = epochs["targetleft", "targetright"]
    X = targetepochs.get_data()  
    y = targetepochs.events[:, 2]  # target: auditory left vs visual left
    y[y<5] = 1     # relabel the left and right conditions as 1 and 2
    y[y>5] = 2
    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    targetu = np.mean(scores, axis=0)

    newevents, _ = mne.events_from_annotations(filtered, event_id=legaltriggers)

    for i in range(1,len(events)):
        if events[i,2]==4:
            if events[i+1,2]==2:
                newevents[i+1,2] = 11    #congruent left
            if events[i+1,2]==3:
                newevents[i+1,2] = 11    #congruent left
            if events[i+1,2]==6:
                newevents[i+1,2] = 14    #incongruent right   
            if events[i+1,2]==7:
                newevents[i+1,2] = 14    #incongruent right
        if events[i,2]==5:
             if events[i+1,2]==2:
                 newevents[i+1,2] = 13    #incongruent left
             if events[i+1,2]==3:
                 newevents[i+1,2] = 13    #incongruent left
             if events[i+1,2]==6:
                 newevents[i+1,2] = 12    #congruent right
             if events[i+1,2]==7:
                 newevents[i+1,2] = 12    #congruent right

    event_dict2 = {
         "congruent/left": 11,
         "congruent/right": 12,
         "incongruent/left": 13,
         "incongruent/right": 14     
     }

    epochsCI = mne.Epochs(
        filtered,
        newevents,
        event_id=event_dict2,
        tmin=-0.2,
        tmax=1,
        reject=reject_criteria,
        preload=True,
    )

    epochsCI.equalize_event_counts(["congruent", "incongruent"])
    targetepochs = epochsCI["congruent", "incongruent"]
    X = targetepochs.get_data()  
    y = targetepochs.events[:, 2]  # target: auditory left vs visual left
    y[y<13] = 1     # relabel the left and right conditions as 1 and 2
    y[y>12] = 2

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    congincong = np.mean(scores, axis=0)

    epochsCI.equalize_event_counts(["left", "right"])
    targetepochs = epochsCI["left", "right"]
    X = targetepochs.get_data() 
    y = targetepochs.events[:, 2]  # target: auditory left vs visual left
    y[y==13] = 1     # relabel the left and right conditions as 1 and 2
    y[y==14] = 2

    i1 = np.array(np.where(y==11))
    i2 = np.random.permutation(i1.shape[1])
    y[i1[0,i2[0:49]]] = 1
    i1 = np.array(np.where(y==12))
    i2 = np.random.permutation(i1.shape[1])
    y[i1[0,i2[0:49]]] = 2
    i1 = np.array(np.where(y<10))
    y = y[i1[0,:]]
    X = X[i1[0,:],:,:]

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    targetb = np.mean(scores, axis=0)

    # export decoding results to csv text files
    filename = "local/processedE1/S" + str(sublist[subj]) + "_decoding.csv"

    times = pd.Series(range(-200,1001))
    cuedecode = pd.Series(cuedecode)
    targetu = pd.Series(targetu)
    congincong = pd.Series(congincong)
    targetb = pd.Series(targetb)

    output = pd.concat([times,cuedecode,targetu,congincong,targetb],axis=1)
    output.columns = ['Time','Cue','TargetU','CongIncong','TargetB']

    output.to_csv(filename,index=False)

    filename = "local/processedE1/S" + str(sublist[subj]) + "_cueleft.csv"
    ERP = epochs["cueleft"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE1/S" + str(sublist[subj]) + "_cueright.csv"
    ERP = epochs["cueright"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE1/S" + str(sublist[subj]) + "_targetleft.csv"
    ERP = epochs["targetleft"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE1/S" + str(sublist[subj]) + "_targetright.csv"
    ERP = epochs["targetright"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE1/S" + str(sublist[subj]) + "_events.csv"
    dataout = np.asarray(events)
    np.savetxt(filename, dataout, delimiter=",")

```

```{python ProcessExpt2}

# code to import raw EEG data and process

if r.processdata==2:

  import numpy as np
  import matplotlib.pyplot as plt
  import pandas as pd
  import mne
  from mne.decoding import (
      SlidingEstimator,
      GeneralizingEstimator,
      Scaler,
      cross_val_multiscore,
      LinearModel,
      get_coef,
      Vectorizer,
      CSP
  )

  from sklearn.pipeline import make_pipeline
  from sklearn.preprocessing import StandardScaler
  from sklearn.linear_model import LogisticRegression

  ANT_montage = mne.channels.make_standard_montage("standard_1020")
  reject_criteria = None   #dict(eeg=150e-6)

  clf = make_pipeline(StandardScaler(), LogisticRegression(solver="liblinear"))
  time_decod = SlidingEstimator(clf, n_jobs=10, scoring="roc_auc", verbose=True)

  legaltriggers = {
    "1": 1,
    "11": 2,
    "12": 3,
    "101": 4,
    "102": 5,
    "21": 6,
    "22": 7,
    "31": 8,
    "32": 9,
    "2": 10
    }
  sublist = list(range(201,202)) + list(range(204,213)) + list(range(214,226))

  for subj in range(0,len(sublist)):
    filename = "local/rawdata/S" + str(sublist[subj]) + ".set"

    raw = mne.io.read_raw_eeglab(filename,preload=True)
    raw.drop_channels(["HEOG", "VEOG", "M1", "M2"])
    raw.set_montage(ANT_montage)

    filtered = raw.filter(0.2,30)

    ica = mne.preprocessing.ICA(n_components=20, random_state=97, max_iter=800)
    ica.fit(filtered)
    ica.exclude = [1, 2]  # details on how we picked these are omitted here
    ica.apply(filtered)

    events, _ = mne.events_from_annotations(filtered, event_id=legaltriggers)

    event_dict = {
        "cueleft": 4,
        "cueright": 5,
        "targetleft/top": 2,
        "targetleft/bottom": 3,
        "targetright/top": 6,
        "targetright/bottom": 7,
        "responsetop": 8,
        "responsebottom": 9,
        "nonstimming": 1,
        "stimming": 10
    }

    epochs = mne.Epochs(
        filtered,
        events,
        event_id=event_dict,
        tmin=-0.2,
        tmax=1,
        reject=reject_criteria,
        preload=True,
    )

    epochs.equalize_event_counts(["cueleft", "cueright"])
    cueepochs = epochs["cueleft", "cueright"]
    X = cueepochs.get_data()  # EEG signals: n_epochs, n_meg_channels, n_times
    y = cueepochs.events[:, 2]  # target: auditory left vs visual left

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    cuedecode = np.mean(scores, axis=0)

    epochs.equalize_event_counts(["targetleft", "targetright"])
    targetepochs = epochs["targetleft", "targetright"]
    X = targetepochs.get_data()  
    y = targetepochs.events[:, 2]  # target: auditory left vs visual left
    y[y<5] = 1     # relabel the left and right conditions as 1 and 2
    y[y>5] = 2
    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    targetu = np.mean(scores, axis=0)

    newevents, _ = mne.events_from_annotations(filtered, event_id=legaltriggers)

    for i in range(1,len(events)):
        if events[i,2]==4:
            if events[i+1,2]==2:
                newevents[i+1,2] = 11    #congruent left
            if events[i+1,2]==3:
                newevents[i+1,2] = 11    #congruent left
            if events[i+1,2]==6:
                newevents[i+1,2] = 14    #incongruent right   
            if events[i+1,2]==7:
                newevents[i+1,2] = 14    #incongruent right
        if events[i,2]==5:
             if events[i+1,2]==2:
                 newevents[i+1,2] = 13    #incongruent left
             if events[i+1,2]==3:
                 newevents[i+1,2] = 13    #incongruent left
             if events[i+1,2]==6:
                 newevents[i+1,2] = 12    #congruent right
             if events[i+1,2]==7:
                 newevents[i+1,2] = 12    #congruent right

    event_dict2 = {
         "congruent/left": 11,
         "congruent/right": 12,
         "incongruent/left": 13,
         "incongruent/right": 14     
     }

    epochsCI = mne.Epochs(
        filtered,
        newevents,
        event_id=event_dict2,
        tmin=-0.2,
        tmax=1,
        reject=reject_criteria,
        preload=True,
    )

    epochsCI.equalize_event_counts(["congruent", "incongruent"])
    targetepochs = epochsCI["congruent", "incongruent"]
    X = targetepochs.get_data()  
    y = targetepochs.events[:, 2]  # target: auditory left vs visual left
    y[y<13] = 1     # relabel the left and right conditions as 1 and 2
    y[y>12] = 2

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    congincong = np.mean(scores, axis=0)

    epochsCI.equalize_event_counts(["left", "right"])
    targetepochs = epochsCI["left", "right"]
    X = targetepochs.get_data() 
    y = targetepochs.events[:, 2]  # target: auditory left vs visual left
    y[y==13] = 1     # relabel the left and right conditions as 1 and 2
    y[y==14] = 2

    i1 = np.array(np.where(y==11))
    i2 = np.random.permutation(i1.shape[1])
    y[i1[0,i2[0:49]]] = 1
    i1 = np.array(np.where(y==12))
    i2 = np.random.permutation(i1.shape[1])
    y[i1[0,i2[0:49]]] = 2
    i1 = np.array(np.where(y<10))
    y = y[i1[0,:]]
    X = X[i1[0,:],:,:]

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    targetb = np.mean(scores, axis=0)

    filename = "local/processedE2/S" + str(sublist[subj]) + "_cueleft.csv"
    ERP = epochs["cueleft"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE2/S" + str(sublist[subj]) + "_cueright.csv"
    ERP = epochs["cueright"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE2/S" + str(sublist[subj]) + "_targetleft.csv"
    ERP = epochs["targetleft"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE2/S" + str(sublist[subj]) + "_targetright.csv"
    ERP = epochs["targetright"].average()
    dataout = np.transpose(np.asarray(ERP.data))*1000000
    np.savetxt(filename, dataout, delimiter=",")

    filename = "local/processedE2/S" + str(sublist[subj]) + "_events.csv"
    dataout = np.asarray(events)
    np.savetxt(filename, dataout, delimiter=",")

    newevents2, _ = mne.events_from_annotations(filtered, event_id=legaltriggers)
    stimcond = 0

    for i in range(0,len(events)):
        if events[i,2] == 1: stimcond = 1 
        if events[i,2] == 10: stimcond = 2 
        if events[i,2] in range(2,10): newevents2[i,2] = events[i,2] + (100*stimcond)

    event_dict3 = {
        "nostimcueleft": 104,
        "nostimcueright": 105,
        "nostimtargetleft/top": 102,
        "nostimtargetleft/bottom": 103,
        "nostimtargetright/top": 106,
        "nostimtargetright/bottom": 107,
        "stimcueleft": 204,
        "stimcueright": 205,
        "stimtargetleft/top": 202,
        "stimtargetleft/bottom": 203,
        "stimtargetright/top": 206,
        "stimtargetright/bottom": 207
    }

    splitepochs = mne.Epochs(
        filtered,
        newevents2,
        event_id=event_dict3,
        tmin=-0.2,
        tmax=1,
        reject=reject_criteria,
        preload=True,
    )

    splitepochs.equalize_event_counts(["nostimcueleft", "nostimcueright"])
    cueepochs2 = splitepochs["nostimcueleft", "nostimcueright"]
    X = cueepochs2.get_data()  # EEG signals: n_epochs, n_meg_channels, n_times
    y = cueepochs2.events[:, 2]-100  # target: auditory left vs visual left

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    cuedecodeNostim = np.mean(scores, axis=0)

    splitepochs.equalize_event_counts(["nostimtargetleft", "nostimtargetright"])
    targetepochs2 = splitepochs["nostimtargetleft", "nostimtargetright"]
    X = targetepochs2.get_data()
    y = targetepochs2.events[:, 2]  # target: auditory left vs visual left
    y[y<105] = 1     # relabel the left and right conditions as 1 and 2
    y[y>105] = 2
    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    targetuNostim = np.mean(scores, axis=0)

    splitepochs.equalize_event_counts(["stimcueleft", "stimcueright"])
    cueepochs3 = splitepochs["stimcueleft", "stimcueright"]
    X = cueepochs3.get_data()  # EEG signals: n_epochs, n_meg_channels, n_times
    y = cueepochs3.events[:, 2]-200  # target: auditory left vs visual left

    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    cuedecodeStim = np.mean(scores, axis=0)

    splitepochs.equalize_event_counts(["stimtargetleft", "stimtargetright"])
    targetepochs3 = splitepochs["stimtargetleft", "stimtargetright"]
    X = targetepochs3.get_data()
    y = targetepochs3.events[:, 2]  # target: auditory left vs visual left
    y[y<205] = 1     # relabel the left and right conditions as 1 and 2
    y[y>205] = 2
    scores = cross_val_multiscore(time_decod, X, y, cv=10, n_jobs=10)
    targetuStim = np.mean(scores, axis=0)

    # export decoding results to csv text files
    filename = "local/processedE2/S" + str(sublist[subj]) + "_decoding.csv"

    times = pd.Series(range(-200,1001))
    cuedecode = pd.Series(cuedecode)
    targetu = pd.Series(targetu)
    congincong = pd.Series(congincong)
    targetb = pd.Series(targetb)
    cuedecodeN = pd.Series(cuedecodeNostim)
    cuedecodeS = pd.Series(cuedecodeStim)
    targetuN = pd.Series(targetuNostim)
    targetuS = pd.Series(targetuStim)

    output = pd.concat([times,cuedecode,targetu,congincong,targetb,cuedecodeN,cuedecodeS,targetuN,targetuS],axis=1)
    output.columns = ['Time','Cue','TargetU','CongIncong','TargetB','CueNostim','CueStim','TargetUNostim','TargetUStim']

    output.to_csv(filename,index=False)

```

```{r collateprocessed1}
#| warning: false

if (processdata>0){

  d <- dir('local/processedE1',pattern='*_decoding.csv',full.names=TRUE)

  allCue <- matrix(0,nrow=1201,ncol=length(d))
  allTargetU <- matrix(0,nrow=1201,ncol=length(d))
  allCongIncong <- matrix(0,nrow=1201,ncol=length(d))
  allTargetB <- matrix(0,nrow=1201,ncol=length(d))
  for (s in 1:length(d)){
    decodingdata <- read.csv(d[s])
    allCue[,s] <- decodingdata$Cue
    allTargetU[,s] <- decodingdata$TargetU
    allCongIncong[,s] <- decodingdata$CongIncong
    allTargetB[,s] <- decodingdata$TargetB
  }
  decodingBFs <- matrix(0,nrow=1201,ncol=4)
  for (t in 1:nrow(allCue)){
    decodingBFs[t,1] <- extractBF(ttestBF(allCue[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs[t,2] <- extractBF(ttestBF(allTargetU[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs[t,3] <- extractBF(ttestBF(allCongIncong[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs[t,4] <- extractBF(ttestBF(allTargetB[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
  }

  meandecoding <- decodingdata
  meandecoding$Cue <- rowMeans(allCue)
  meandecoding$TargetU <- rowMeans(allTargetU)
  meandecoding$CongIncong <- rowMeans(allCongIncong)
  meandecoding$TargetB <- rowMeans(allTargetB)

  CIdecoding <- decodingdata
  CIdecoding$Cue <- apply(allCue,1,sd)/sqrt(ncol(allCue))*1.96
  CIdecoding$TargetU <- apply(allTargetU,1,sd)/sqrt(ncol(allTargetU))*1.96
  CIdecoding$CongIncong <- apply(allCongIncong,1,sd)/sqrt(ncol(allCongIncong))*1.96
  CIdecoding$TargetB <- apply(allTargetB,1,sd)/sqrt(ncol(allTargetB))*1.96

  d <- dir('local/processedE1',pattern='*_cueleft.csv',full.names=TRUE)
  allcueleft <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    allcueleft[s,,] <- as.matrix(ERPdata)
  }
  d <- dir('local/processedE1',pattern='*_cueright.csv',full.names=TRUE)
  allcueright <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    allcueright[s,,] <- as.matrix(ERPdata)
  }
  d <- dir('local/processedE1',pattern='*_targetleft.csv',full.names=TRUE)
  alltargetleft <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    alltargetleft[s,,] <- as.matrix(ERPdata)
  }
  d <- dir('local/processedE1',pattern='*_targetright.csv',full.names=TRUE)
  alltargetright <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    alltargetright[s,,] <- as.matrix(ERPdata)
  }

  meancueleft <- getcleanmean(allcueleft)
  meancueright <- getcleanmean(allcueright)
  meantargetleft <- getcleanmean(alltargetleft)
  meantargetright <- getcleanmean(alltargetright)

  CIcueleft <- getcleanCI(allcueleft)
  CIcueright <- getcleanCI(allcueright)
  CItargetleft <- getcleanCI(alltargetleft)
  CItargetright <- getcleanCI(alltargetright)

  save(file='local/Exp1Group.RData',list=c('meandecoding','CIdecoding','meancueleft','CIcueleft','meancueright','CIcueright','meantargetleft','CItargetleft','meantargetright','CItargetright','decodingBFs'))

}

```

```{r collateprocessed2}
#| warning: false

if (processdata>0){

  d <- dir('local/processedE2',pattern='*_decoding.csv',full.names=TRUE)
  d2 <- dir('local/processedE2',pattern='*_decoding.csv',full.names=FALSE)
  sublist <- as.numeric(gsub("\\D", "", d2))

  stimindices <- which(sublist %in% stimlist)
  nostimindices <- which(sublist %in% nostimlist)

  allCue <- matrix(0,nrow=1201,ncol=length(d))
  allTargetU <- matrix(0,nrow=1201,ncol=length(d))
  allCongIncong <- matrix(0,nrow=1201,ncol=length(d))
  allTargetB <- matrix(0,nrow=1201,ncol=length(d))
  allCueNostim <- matrix(0,nrow=1201,ncol=length(d))
  allCueStim <- matrix(0,nrow=1201,ncol=length(d))
  for (s in 1:length(d)){
    decodingdata <- read.csv(d[s])
    allCue[,s] <- decodingdata$Cue
    allTargetU[,s] <- decodingdata$TargetU
    allCongIncong[,s] <- decodingdata$CongIncong
    allTargetB[,s] <- decodingdata$TargetB
    allCueNostim[,s] <- decodingdata$CueNostim
    allCueStim[,s] <- decodingdata$CueStim
  }
    decodingBFs2 <- matrix(0,nrow=1201,ncol=7)
  for (t in 1:nrow(allCue)){
    decodingBFs2[t,1] <- extractBF(ttestBF(allCue[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,2] <- extractBF(ttestBF(allTargetU[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,3] <- extractBF(ttestBF(allCongIncong[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,4] <- extractBF(ttestBF(allTargetB[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,5] <- extractBF(ttestBF(allCueNostim[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,6] <- extractBF(ttestBF(allCueStim[t,],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,7] <- extractBF(ttestBF(allCueStim[t,],allCueNostim[t,],paired=TRUE))$bf
  }

    d <- dir('local/processedE2',pattern='*_cueleft.csv',full.names=TRUE)
  allcueleft <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    allcueleft[s,,] <- as.matrix(ERPdata)
  }
  d <- dir('local/processedE2',pattern='*_cueright.csv',full.names=TRUE)
  allcueright <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    allcueright[s,,] <- as.matrix(ERPdata)
  }
  d <- dir('local/processedE2',pattern='*_targetleft.csv',full.names=TRUE)
  alltargetleft <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    alltargetleft[s,,] <- as.matrix(ERPdata)
  }
  d <- dir('local/processedE2',pattern='*_targetright.csv',full.names=TRUE)
  alltargetright <- array(0,dim=c(length(d),1201,62))
  for (s in 1:length(d)){
    ERPdata <- read.csv(d[s],header=FALSE)
    alltargetright[s,,] <- as.matrix(ERPdata)
  }

  meandecoding <- decodingdata
  meandecoding$Cue <- rowMeans(allCue)
  meandecoding$TargetU <- rowMeans(allTargetU)
  meandecoding$CongIncong <- rowMeans(allCongIncong)
  meandecoding$TargetB <- rowMeans(allTargetB)
  meandecoding$CueNostim <- rowMeans(allCueNostim)
  meandecoding$CueStim <- rowMeans(allCueStim)

  CIdecoding <- decodingdata
  CIdecoding$Cue <- apply(allCue,1,sd)/sqrt(ncol(allCue))*1.96
  CIdecoding$TargetU <- apply(allTargetU,1,sd)/sqrt(ncol(allTargetU))*1.96
  CIdecoding$CongIncong <- apply(allCongIncong,1,sd)/sqrt(ncol(allCongIncong))*1.96
  CIdecoding$TargetB <- apply(allTargetB,1,sd)/sqrt(ncol(allTargetB))*1.96
  CIdecoding$CueNoStim <- apply(allCueNostim,1,sd)/sqrt(ncol(allCueNostim))*1.96
  CIdecoding$CueStim <- apply(allCueStim,1,sd)/sqrt(ncol(allCueStim))*1.96

  meancueleft <- getcleanmean(allcueleft)
  meancueright <- getcleanmean(allcueright)
  meantargetleft <- getcleanmean(alltargetleft)
  meantargetright <- getcleanmean(alltargetright)

  CIcueleft <- getcleanCI(allcueleft)
  CIcueright <- getcleanCI(allcueright)
  CItargetleft <- getcleanCI(alltargetleft)
  CItargetright <- getcleanCI(alltargetright)

  save(file='local/Exp2Group.RData',list=c('meandecoding','CIdecoding','meancueleft','CIcueleft','meancueright','CIcueright','meantargetleft','CItargetleft','meantargetright','CItargetright','decodingBFs2'))

  meandecoding <- decodingdata
  meandecoding$Cue <- rowMeans(allCue[,stimindices])
  meandecoding$TargetU <- rowMeans(allTargetU[,stimindices])
  meandecoding$CongIncong <- rowMeans(allCongIncong[,stimindices])
  meandecoding$TargetB <- rowMeans(allTargetB[,stimindices])
  meandecoding$CueNostim <- rowMeans(allCueNostim[,stimindices])
  meandecoding$CueStim <- rowMeans(allCueStim[,stimindices])

      decodingBFs2 <- matrix(0,nrow=1201,ncol=7)
      groupcomp <- matrix(0,nrow=1201,ncol=2)
  for (t in 1:nrow(allCue)){
    decodingBFs2[t,1] <- extractBF(ttestBF(allCue[t,stimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,2] <- extractBF(ttestBF(allTargetU[t,stimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,3] <- extractBF(ttestBF(allCongIncong[t,stimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,4] <- extractBF(ttestBF(allTargetB[t,stimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,5] <- extractBF(ttestBF(allCueNostim[t,stimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,6] <- extractBF(ttestBF(allCueStim[t,stimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,7] <- extractBF(ttestBF(allCueStim[t,stimindices],allCueNostim[t,stimindices],paired=TRUE))$bf

    groupcomp[t,1] <- extractBF(ttestBF(allCue[t,nostimindices],allCue[t,stimindices],paired=FALSE))$bf
    groupcomp[t,2] <- extractBF(ttestBF(allCongIncong[t,nostimindices],allCongIncong[t,stimindices],paired=FALSE))$bf
  }

  CIdecoding <- decodingdata
  CIdecoding$Cue <- apply(allCue[,stimindices],1,sd)/sqrt(ncol(allCue[,stimindices]))*1.96
  CIdecoding$TargetU <- apply(allTargetU[,stimindices],1,sd)/sqrt(ncol(allTargetU[,stimindices]))*1.96
  CIdecoding$CongIncong <- apply(allCongIncong[,stimindices],1,sd)/sqrt(ncol(allCongIncong[,stimindices]))*1.96
  CIdecoding$TargetB <- apply(allTargetB[,stimindices],1,sd)/sqrt(ncol(allTargetB[,stimindices]))*1.96
  CIdecoding$CueNoStim <- apply(allCueNostim[,stimindices],1,sd)/sqrt(ncol(allCueNostim[,stimindices]))*1.96
  CIdecoding$CueStim <- apply(allCueStim[,stimindices],1,sd)/sqrt(ncol(allCueStim[,stimindices]))*1.96

  meancueleft <- getcleanmean(allcueleft[stimindices,,])
  meancueright <- getcleanmean(allcueright[stimindices,,])
  meantargetleft <- getcleanmean(alltargetleft[stimindices,,])
  meantargetright <- getcleanmean(alltargetright[stimindices,,])

  CIcueleft <- getcleanCI(allcueleft[stimindices,,])
  CIcueright <- getcleanCI(allcueright[stimindices,,])
  CItargetleft <- getcleanCI(alltargetleft[stimindices,,])
  CItargetright <- getcleanCI(alltargetright[stimindices,,])

  save(file='local/Exp2GroupStim.RData',list=c('meandecoding','CIdecoding','meancueleft','CIcueleft','meancueright','CIcueright','meantargetleft','CItargetleft','meantargetright','CItargetright', 'decodingBFs2','groupcomp'))

  meandecoding <- decodingdata
  meandecoding$Cue <- rowMeans(allCue[,nostimindices])
  meandecoding$TargetU <- rowMeans(allTargetU[,nostimindices])
  meandecoding$CongIncong <- rowMeans(allCongIncong[,nostimindices])
  meandecoding$TargetB <- rowMeans(allTargetB[,nostimindices])
  meandecoding$CueNostim <- rowMeans(allCueNostim[,nostimindices])
  meandecoding$CueStim <- rowMeans(allCueStim[,nostimindices])

        decodingBFs2 <- matrix(0,nrow=1201,ncol=7)
  for (t in 1:nrow(allCue)){
    decodingBFs2[t,1] <- extractBF(ttestBF(allCue[t,nostimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,2] <- extractBF(ttestBF(allTargetU[t,nostimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,3] <- extractBF(ttestBF(allCongIncong[t,nostimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,4] <- extractBF(ttestBF(allTargetB[t,nostimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,5] <- extractBF(ttestBF(allCueNostim[t,nostimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,6] <- extractBF(ttestBF(allCueStim[t,nostimindices],mu=0.5,nullInterval=c(0.5,Inf)))$bf[1]
    decodingBFs2[t,7] <- extractBF(ttestBF(allCueStim[t,nostimindices],allCueNostim[t,nostimindices],paired=TRUE))$bf
  }
  CIdecoding <- decodingdata
  CIdecoding$Cue <- apply(allCue[,nostimindices],1,sd)/sqrt(ncol(allCue[,nostimindices]))*1.96
  CIdecoding$TargetU <- apply(allTargetU[,nostimindices],1,sd)/sqrt(ncol(allTargetU[,nostimindices]))*1.96
  CIdecoding$CongIncong <- apply(allCongIncong[,nostimindices],1,sd)/sqrt(ncol(allCongIncong[,nostimindices]))*1.96
  CIdecoding$TargetB <- apply(allTargetB[,nostimindices],1,sd)/sqrt(ncol(allTargetB[,nostimindices]))*1.96
  CIdecoding$CueNoStim <- apply(allCueNostim[,nostimindices],1,sd)/sqrt(ncol(allCueNostim[,nostimindices]))*1.96
  CIdecoding$CueStim <- apply(allCueStim[,nostimindices],1,sd)/sqrt(ncol(allCueStim[,nostimindices]))*1.96

  meancueleft <- getcleanmean(allcueleft[nostimindices,,])
  meancueright <- getcleanmean(allcueright[nostimindices,,])
  meantargetleft <- getcleanmean(alltargetleft[nostimindices,,])
  meantargetright <- getcleanmean(alltargetright[nostimindices,,])

  CIcueleft <- getcleanCI(allcueleft[nostimindices,,])
  CIcueright <- getcleanCI(allcueright[nostimindices,,])
  CItargetleft <- getcleanCI(alltargetleft[nostimindices,,])
  CItargetright <- getcleanCI(alltargetright[nostimindices,,])

  save(file='local/Exp2GroupNoStim.RData',list=c('meandecoding','CIdecoding','meancueleft','CIcueleft','meancueright','CIcueright','meantargetleft','CItargetleft','meantargetright','CItargetright','decodingBFs2','groupcomp'))

}

```

```{r analyseresponses1}
#| warning: false

if (processdata>0){

  d <- dir('local/processedE1',pattern='_events.csv',full.names=TRUE)

  allRT <- matrix(0, nrow=length(d), 2)
  allaccuracy <- matrix(0, nrow=length(d), 2)

  for (s in 1:length(d)){
  respdata <- read.csv(d[s],header=FALSE)

  cueside <- NULL
  targetside <- NULL
  targetpos <- NULL
  iscorrect <- NULL
  RT <- NULL
  trialcount <- 0
  listpos <- 0

  while (listpos<nrow(respdata)){
    listpos <- listpos + 1
    if (sum(respdata[listpos,3]==c(4,5))){
      trialcount <- trialcount + 1
      cueside[trialcount] <- respdata[listpos,3]-3
    }
    if (sum(respdata[listpos,3]==c(2,3,6,7))){
      targetonset <- respdata[listpos,1]
    }        
    if (sum(respdata[listpos,3]==c(2,3))){
      targetside[trialcount] <- 1
    }    
    if (sum(respdata[listpos,3]==c(6,7))){
      targetside[trialcount] <- 2
    }   
    if (sum(respdata[listpos,3]==c(2,6))){
      targetpos[trialcount] <- 1
    }    
    if (sum(respdata[listpos,3]==c(3,7))){
      targetpos[trialcount] <- 2
    }

    if (sum(respdata[listpos,3]==c(8,9))){
      if ((respdata[listpos,3]==8) & (targetpos[trialcount]==1)){iscorrect[trialcount] <- 0}
      if ((respdata[listpos,3]==8) & (targetpos[trialcount]==2)){iscorrect[trialcount] <- 1}
      if ((respdata[listpos,3]==9) & (targetpos[trialcount]==1)){iscorrect[trialcount] <- 1}
      if ((respdata[listpos,3]==9) & (targetpos[trialcount]==2)){iscorrect[trialcount] <- 0}
      RT[trialcount] <- respdata[listpos,1] - targetonset
    }

  }

  congincong <- cueside==targetside
  subjresps <- data.frame(cueside,targetside,congincong,targetpos,iscorrect,RT)

  allcong <- subset(subjresps,congincong==TRUE)
  allaccuracy[s,1] <- mean(allcong$iscorrect)
  allRT[s,1] <- mean(20*log10(allcong$RT))
  allincong <- subset(subjresps,congincong==FALSE)
  allaccuracy[s,2] <- mean(allincong$iscorrect)
  allRT[s,2] <- mean(20*log10(allincong$RT))
  }

  save(file='local/Expt1resps.RData',list=c('allRT','allaccuracy'))
}

load('local/Expt1resps.RData')

ACCtest <- t.test(allaccuracy[,1],allaccuracy[,2],paired=TRUE)
temp <- allaccuracy[,1]-allaccuracy[,2]
ACCd <- abs(mean(temp))/sd(temp)
accdiff <- 100*(mean(allaccuracy[,1])-mean(allaccuracy[,2]))
ACCBF <- ttestBF(allaccuracy[,1],allaccuracy[,2],paired=TRUE)

RTtest <- t.test(allRT[,1],allRT[,2],paired=TRUE)
temp <- allRT[,1]-allRT[,2]
RTd <- abs(mean(temp))/sd(temp)
RTdiff <- 10^(mean(allRT[,2])/20)-10^(mean(allRT[,1])/20)
RTBF <- ttestBF(allRT[,1],allRT[,2],paired=TRUE)

```

```{r analyseresponses2}
#| warning: false

if (processdata>0){

  d <- dir('local/processedE2',pattern='_events.csv',full.names=TRUE)
  d2 <- dir('local/processedE2',pattern='_events.csv',full.names=FALSE)

  allRT <- matrix(0, nrow=length(d), 4)
  allaccuracy <- matrix(0, nrow=length(d), 4)
  reportsstimming <- 0*(1:length(d))
  subjIDs <- parse_number(d2)

  for (s in 1:length(d)){
  respdata <- read.csv(d[s],header=FALSE)

  if (sum(subjIDs[s]==stimlist)){reportsstimming[s] <- 1}

  cueside <- NULL
  targetside <- NULL
  targetpos <- NULL
  iscorrect <- NULL
  RT <- NULL
  stimcond <- NULL
  trialcount <- 0
  listpos <- 0
  stimblock <- 0

  while (listpos<nrow(respdata)){
    listpos <- listpos + 1
    if (respdata[listpos,3]==1){stimblock <- 1}
    if (respdata[listpos,3]==10){stimblock <- 2}

    if (sum(respdata[listpos,3]==c(4,5))){
      trialcount <- trialcount + 1
      cueside[trialcount] <- respdata[listpos,3]-3
    }
    if (sum(respdata[listpos,3]==c(2,3,6,7))){
      targetonset <- respdata[listpos,1]
    }        
    if (sum(respdata[listpos,3]==c(2,3))){
      targetside[trialcount] <- 1
    }    
    if (sum(respdata[listpos,3]==c(6,7))){
      targetside[trialcount] <- 2
    }   
    if (sum(respdata[listpos,3]==c(2,6))){
      targetpos[trialcount] <- 1
    }    
    if (sum(respdata[listpos,3]==c(3,7))){
      targetpos[trialcount] <- 2
    }

    stimcond[trialcount] <- stimblock

    if (sum(respdata[listpos,3]==c(8,9))){
      if ((respdata[listpos,3]==8) & (targetpos[trialcount]==1)){iscorrect[trialcount] <- 0}
      if ((respdata[listpos,3]==8) & (targetpos[trialcount]==2)){iscorrect[trialcount] <- 1}
      if ((respdata[listpos,3]==9) & (targetpos[trialcount]==1)){iscorrect[trialcount] <- 1}
      if ((respdata[listpos,3]==9) & (targetpos[trialcount]==2)){iscorrect[trialcount] <- 0}
      RT[trialcount] <- respdata[listpos,1] - targetonset
    }

  }

  congincong <- cueside==targetside
  subjresps <- data.frame(stimcond,cueside,targetside,congincong,targetpos,iscorrect,RT)

  temp <- subset(subjresps,stimcond==1)
  allcong <- subset(temp,congincong==TRUE)
  allaccuracy[s,1] <- mean(allcong$iscorrect)
  allRT[s,1] <- mean(20*log10(allcong$RT))
  allincong <- subset(temp,congincong==FALSE)
  allaccuracy[s,2] <- mean(allincong$iscorrect)
  allRT[s,2] <- mean(20*log10(allincong$RT))

  temp <- subset(subjresps,stimcond==2)
  allcong <- subset(temp,congincong==TRUE)
  allaccuracy[s,3] <- mean(allcong$iscorrect)
  allRT[s,3] <- mean(20*log10(allcong$RT))
  allincong <- subset(temp,congincong==FALSE)
  allaccuracy[s,4] <- mean(allincong$iscorrect)
  allRT[s,4] <- mean(20*log10(allincong$RT))

  }

  save(file='local/Expt2resps.RData',list=c('allRT','allaccuracy','reportsstimming'))
}

load('local/Expt2resps.RData')

# restructure for 2x2 ANOVA

accuracy <- as.vector(allaccuracy)
participant <- as.factor(rep(1:nrow(allaccuracy),4))
stimcond <- as.factor(rep(1:2,each=(2*nrow(allaccuracy))))
congruent <- as.factor(rep(1:2,each=nrow(allaccuracy),2))
accuracyaov <- data.frame(accuracy,participant,stimcond,congruent)
accAOVresults <- ezANOVA(accuracyaov, dv=accuracy, wid=participant, within=c(stimcond,congruent))
accAOVBF <- anovaBF(accuracy ~ stimcond * congruent + participant, data=accuracyaov, whichModels='all', whichRandom = 'participant')

RT <- as.vector(allRT)
participant <- as.factor(rep(1:nrow(allRT),4))
stimcond <- as.factor(rep(1:2,each=(2*nrow(allRT))))
congruent <- as.factor(rep(1:2,each=nrow(allRT),2))
RTaov <- data.frame(RT,participant,stimcond,congruent)
RTAOVresults <- ezANOVA(RTaov, dv=RT, wid=participant, within=c(stimcond,congruent))
RTAOVBF <- anovaBF(RT ~ stimcond * congruent + participant, data=RTaov, whichModels='all', whichRandom = 'participant')

stimaccuracy <- allaccuracy[which(reportsstimming>0),]
accuracy <- as.vector(stimaccuracy)
participant <- as.factor(rep(1:nrow(stimaccuracy),4))
stimcond <- as.factor(rep(1:2,each=(2*nrow(stimaccuracy))))
congruent <- as.factor(rep(1:2,each=nrow(stimaccuracy),2))
accuracyaovSTIM <- data.frame(accuracy,participant,stimcond,congruent)
accAOVresultsSTIM <- ezANOVA(accuracyaovSTIM, dv=accuracy, wid=participant, within=c(stimcond,congruent))
accAOVBFSTIM <- anovaBF(accuracy ~ stimcond * congruent + participant, data=accuracyaovSTIM, whichModels='all', whichRandom = 'participant')

stimRT <- allRT[which(reportsstimming>0),]
RT <- as.vector(stimRT)
participant <- as.factor(rep(1:nrow(stimRT),4))
stimcond <- as.factor(rep(1:2,each=(2*nrow(stimRT))))
congruent <- as.factor(rep(1:2,each=nrow(stimRT),2))
RTaovSTIM <- data.frame(RT,participant,stimcond,congruent)
RTAOVresultsSTIM <- ezANOVA(RTaovSTIM, dv=RT, wid=participant, within=c(stimcond,congruent))
RTAOVBFSTIM <- anovaBF(RT ~ stimcond * congruent + participant, data=RTaovSTIM, whichModels='all', whichRandom = 'participant')

nostimaccuracy <- allaccuracy[which(reportsstimming==0),]
accuracy <- as.vector(nostimaccuracy)
participant <- as.factor(rep(1:nrow(nostimaccuracy),4))
stimcond <- as.factor(rep(1:2,each=(2*nrow(nostimaccuracy))))
congruent <- as.factor(rep(1:2,each=nrow(nostimaccuracy),2))
accuracyaovNOSTIM <- data.frame(accuracy,participant,stimcond,congruent)
accAOVresultsNOSTIM <- ezANOVA(accuracyaovNOSTIM, dv=accuracy, wid=participant, within=c(stimcond,congruent))
accAOVBFnoSTIM <- anovaBF(accuracy ~ stimcond * congruent + participant, data=accuracyaovNOSTIM, whichModels='all', whichRandom = 'participant')

nostimRT <- allRT[which(reportsstimming==0),]
RT <- as.vector(nostimRT)
participant <- as.factor(rep(1:nrow(nostimRT),4))
stimcond <- as.factor(rep(1:2,each=(2*nrow(nostimRT))))
congruent <- as.factor(rep(1:2,each=nrow(nostimRT),2))
RTaovNOSTIM <- data.frame(RT,participant,stimcond,congruent)
RTAOVresultsNOSTIM <- ezANOVA(RTaovNOSTIM, dv=RT, wid=participant, within=c(stimcond,congruent))
RTAOVBFnoSTIM <- anovaBF(RT ~ stimcond * congruent + participant, data=RTaovNOSTIM, whichModels='all', whichRandom = 'participant')

```

```{r makeplotExpt1resps}
#| include: false
#| output: false
#| warning: false

if (processdata>0){

load('local/Expt1resps.RData')

pdf(paste0(figdir,"Exp1resps.pdf"), bg="transparent", height = 5.5, width = 5.5)

plotlims <- c(49.5,58,0.7,1)  
ticklocsx <- 20*log10(seq(300,800,100))    # locations of tick marks on x axis
ticklocsy <- seq(0.7,1,0.1)    # locations of tick marks on y axis
ticklabelsx <- seq(300,800,100)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Reaction time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

coltrans1=rgb(0,0,1,alpha=0.2)
coltrans2=rgb(1,0,0,alpha=0.2)

a <- density(allRT[,1])
a$y <- 0.05*(a$y/max(a$y))
polygon(a$x, a$y+0.7, col=coltrans1,border=NA) 
a <- density(allRT[,2])
a$y <- 0.05*(a$y/max(a$y))
polygon(a$x, a$y+0.7, col=coltrans2,border=NA)

a <- density(allaccuracy[,1])
a$y <- 1.5*(a$y/max(a$y))
polygon(a$y+49.5, a$x, col=coltrans1,border=NA) 
a <- density(allaccuracy[,2])
a$y <- 1.5*(a$y/max(a$y))
polygon(a$y+49.5, a$x, col=coltrans2,border=NA)

for (n in 1:nrow(allRT)){lines(allRT[n,],allaccuracy[n,],col='grey')}
points(allRT[,1],allaccuracy[,1],pch=16,col=flatalpha('blue',0.4))
points(allRT[,2],allaccuracy[,2],pch=16,col=flatalpha('red',0.4))

for (cond in 1:2){
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond])+(1.96*sd(allRT[,cond])/sqrt(nrow(allRT))),mean(allaccuracy[,cond]),angle=90,lwd=2,length=0.05)
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond])-(1.96*sd(allRT[,cond])/sqrt(nrow(allRT))),mean(allaccuracy[,cond]),angle=90,lwd=2,length=0.05)

arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond]),mean(allaccuracy[,cond])+(1.96*sd(allaccuracy[,cond])/sqrt(nrow(allaccuracy))),angle=90,lwd=2,length=0.05)
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond]),mean(allaccuracy[,cond])-(1.96*sd(allaccuracy[,cond])/sqrt(nrow(allaccuracy))),angle=90,lwd=2,length=0.05)
}

points(mean(allRT[,1]),mean(allaccuracy[,1]),pch=21,cex=2,lwd=2,bg='blue')
points(mean(allRT[,2]),mean(allaccuracy[,2]),pch=21,cex=2,lwd=2,bg='red')

legend(50,0.82,c('Congruent','Incongruent'),pch=21,pt.bg=c('blue','red'),pt.cex=2,pt.lwd=2,box.lwd=2)

dev.off()
}

```

```{r makemethodsfig}
#| include: false
#| output: false
#| warning: false

if (processdata>0){

  facestim <- read.bmp('Experiment1code/facel.bmp')
  facestim <- as.raster(matrix(as.numeric(facestim),nrow=600,ncol=600)/255)

  stimenv <- make_soft_window(512,512,0.9)
  gratingstim <- mkgrating(512,8,90,1,1)
  targetL <- gratingstim*stimenv
  targetL[1:256,] <- targetL[1:256,]*0.5
  targetL[257:512,] <- targetL[257:512,]*0.6
  targetL <- (targetL + 1)/2
  targetU <- gratingstim*stimenv
  targetU[1:256,] <- targetU[1:256,]*0.6
  targetU[257:512,] <- targetU[257:512,]*0.5
  targetU <- (targetU + 1)/2

  pdf("Figures/MethodsFigure.pdf", height = 8, width = 19.2)

  par(mar=c(0.1,0.1,0.1,0.1))

  plotlims <- c(0,1.6,1/3,1)  
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])

  polygon(c(0,0,0.2,0.2)+0.1,c(0.8,1,1,0.8),lwd=2,col=rgb(0.5,0.5,0.5))
  rasterImage(facestim,0.25,0.65,0.45,0.85)
  polygon(c(0,0,0.2,0.2)+0.15+0.1,c(0.8,1,1,0.8)-0.15,lwd=2)
  polygon(c(0,0,0.2,0.2)+0.3+0.1,c(0.8,1,1,0.8)-0.3,lwd=2,col=rgb(0.5,0.5,0.5))
  polygon(c(0,0,0.2,0.2)+0.45+0.1,c(0.8,1,1,0.8)-0.45,lwd=2,col=rgb(0.5,0.5,0.5))
  rasterImage(targetL,0.57,0.42,0.63,0.48)

  lines(c(0.2,0.2),c(0.895,0.905))
  lines(c(0.195,0.205),c(0.9,0.9))
  lines(c(0.2,0.2)+0.3,c(0.895,0.905)-0.3)
  lines(c(0.195,0.205)+0.3,c(0.9,0.9)-0.3)
  lines(c(0.2,0.2)+0.45,c(0.895,0.905)-0.45)
  lines(c(0.195,0.205)+0.45,c(0.9,0.9)-0.45)

  arrows(0.05,0.8,0.5,0.35,lwd=4)
  text(0.26,0.56,'Time',adj=0.5,srt=-45,cex=2)
  text(0.32,0.95,'Fixation',pos=4,cex=1.5)
  text(0.47,0.8,'Cue (200ms)',pos=4,cex=1.5)
  text(0.62,0.65,'Fixation (400-600ms)',pos=4,cex=1.5)
  text(0.77,0.5,'Target (200ms)',pos=4,cex=1.5)

  polygon(c(0,0,0.2,0.2)+0.8,c(0.8,1,1,0.8),lwd=2,col=rgb(0.5,0.5,0.5))
  polygon(c(0,0,0.2,0.2)+0.15+0.8,c(0.8,1,1,0.8)-0.15,lwd=2,col=rgb(0.5,0.5,0.5))
  polygon(c(0,0,0.2,0.2)+0.3+0.8,c(0.8,1,1,0.8)-0.3,lwd=2,col=rgb(0.5,0.5,0.5))
  polygon(c(0,0,0.2,0.2)+0.45+0.8,c(0.8,1,1,0.8)-0.45,lwd=2,col=rgb(0.5,0.5,0.5))
  rasterImage(targetU,0.57+0.7,0.42,0.63+0.7,0.48)

  lines(c(0.2,0.2)+0.7,c(0.895,0.905))
  lines(c(0.195,0.205)+0.7,c(0.9,0.9))
  lines(c(0.2,0.2)+0.3+0.7,c(0.895,0.905)-0.3)
  lines(c(0.195,0.205)+0.3+0.7,c(0.9,0.9)-0.3)
  lines(c(0.2,0.2)+0.45+0.7,c(0.895,0.905)-0.45)
  lines(c(0.195,0.205)+0.45+0.7,c(0.9,0.9)-0.45)

  arrows(1.04,0.75,1.06,0.75,lwd=1.5,length=0.1)

  arrows(0.75,0.8,1.2,0.35,lwd=4)
  text(0.96,0.56,'Time',adj=0.5,srt=-45,cex=2)
  text(0.32+0.7,0.95,'Fixation',pos=4,cex=1.5)
  text(0.47+0.7,0.8,'Cue (200ms)',pos=4,cex=1.5)
  text(0.62+0.7,0.65,'Fixation (400-600ms)',pos=4,cex=1.5)
  text(0.77+0.7,0.5,'Target (200ms)',pos=4,cex=1.5)

  text(0.02,0.98,'(a)',cex=4)
  text(0.72,0.98,'(b)',cex=4)

  dev.off()
}

```

```{r makeplotExpt2resps}
#| include: false
#| output: false
#| warning: false

if (processdata>0){

load('local/Expt2resps.RData')

pdf(paste0(figdir,"Exp2resps.pdf"), bg="transparent", height = 5.5, width = 10)

par(mfcol=c(1,2))

plotlims <- c(49.5,58,0.7,1)  
ticklocsx <- 20*log10(seq(300,800,100))    # locations of tick marks on x axis
ticklocsy <- seq(0.7,1,0.1)    # locations of tick marks on y axis
ticklabelsx <- seq(300,800,100)        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Reaction time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main="No stimming", cex.main=1.5)

coltrans1=rgb(0,0,1,alpha=0.2)
coltrans2=rgb(1,0,0,alpha=0.2)

a <- density(allRT[,1])
a$y <- 0.05*(a$y/max(a$y))
polygon(a$x, a$y+0.7, col=coltrans1,border=NA) 
a <- density(allRT[,2])
a$y <- 0.05*(a$y/max(a$y))
polygon(a$x, a$y+0.7, col=coltrans2,border=NA)

a <- density(allaccuracy[,1])
a$y <- 1.5*(a$y/max(a$y))
polygon(a$y+49.5, a$x, col=coltrans1,border=NA) 
a <- density(allaccuracy[,2])
a$y <- 1.5*(a$y/max(a$y))
polygon(a$y+49.5, a$x, col=coltrans2,border=NA)

for (n in 1:nrow(allRT)){lines(allRT[n,1:2],allaccuracy[n,1:2],col='grey')}
points(allRT[,1],allaccuracy[,1],pch=16,col=flatalpha('blue',0.4))
points(allRT[,2],allaccuracy[,2],pch=16,col=flatalpha('red',0.4))

for (cond in 1:2){
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond])+(1.96*sd(allRT[,cond])/sqrt(nrow(allRT))),mean(allaccuracy[,cond]),angle=90,lwd=2,length=0.05)
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond])-(1.96*sd(allRT[,cond])/sqrt(nrow(allRT))),mean(allaccuracy[,cond]),angle=90,lwd=2,length=0.05)

arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond]),mean(allaccuracy[,cond])+(1.96*sd(allaccuracy[,cond])/sqrt(nrow(allaccuracy))),angle=90,lwd=2,length=0.05)
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond]),mean(allaccuracy[,cond])-(1.96*sd(allaccuracy[,cond])/sqrt(nrow(allaccuracy))),angle=90,lwd=2,length=0.05)
}

points(mean(allRT[,1]),mean(allaccuracy[,1]),pch=21,cex=2,lwd=2,bg='blue')
points(mean(allRT[,2]),mean(allaccuracy[,2]),pch=21,cex=2,lwd=2,bg='red')

legend(50,0.82,c('Congruent','Incongruent'),pch=21,pt.bg=c('blue','red'),pt.cex=2,pt.lwd=2,box.lwd=2)

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Reaction time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)
title(main="Stimming", cex.main=1.5)

coltrans1=rgb(0,0,1,alpha=0.2)
coltrans2=rgb(1,0,0,alpha=0.2)

a <- density(allRT[,3])
a$y <- 0.05*(a$y/max(a$y))
polygon(a$x, a$y+0.7, col=coltrans1,border=NA) 
a <- density(allRT[,4])
a$y <- 0.05*(a$y/max(a$y))
polygon(a$x, a$y+0.7, col=coltrans2,border=NA)

a <- density(allaccuracy[,3])
a$y <- 1.5*(a$y/max(a$y))
polygon(a$y+49.5, a$x, col=coltrans1,border=NA) 
a <- density(allaccuracy[,4])
a$y <- 1.5*(a$y/max(a$y))
polygon(a$y+49.5, a$x, col=coltrans2,border=NA)

for (n in 1:nrow(allRT)){lines(allRT[n,3:4],allaccuracy[n,3:4],col='grey')}
points(allRT[,3],allaccuracy[,3],pch=16,col=flatalpha('blue',0.4))
points(allRT[,4],allaccuracy[,4],pch=16,col=flatalpha('red',0.4))

for (cond in 3:4){
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond])+(1.96*sd(allRT[,cond])/sqrt(nrow(allRT))),mean(allaccuracy[,cond]),angle=90,lwd=2,length=0.05)
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond])-(1.96*sd(allRT[,cond])/sqrt(nrow(allRT))),mean(allaccuracy[,cond]),angle=90,lwd=2,length=0.05)

arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond]),mean(allaccuracy[,cond])+(1.96*sd(allaccuracy[,cond])/sqrt(nrow(allaccuracy))),angle=90,lwd=2,length=0.05)
arrows(mean(allRT[,cond]),mean(allaccuracy[,cond]),mean(allRT[,cond]),mean(allaccuracy[,cond])-(1.96*sd(allaccuracy[,cond])/sqrt(nrow(allaccuracy))),angle=90,lwd=2,length=0.05)
}

points(mean(allRT[,3]),mean(allaccuracy[,3]),pch=21,cex=2,lwd=2,bg='blue')
points(mean(allRT[,4]),mean(allaccuracy[,4]),pch=21,cex=2,lwd=2,bg='red')

legend(50,0.82,c('Congruent','Incongruent'),pch=21,pt.bg=c('blue','red'),pt.cex=2,pt.lwd=2,box.lwd=2)

dev.off()
}

```

```{r makeplotExpt1EEG}
#| include: false
#| output: false
#| warning: false

if (processdata>0){

load('local/Exp1Group.RData')

plotlims <- c(-200,1000,0.4,1.1)  
ticklocsx <- seq(-200,1000,200)    # locations of tick marks on x axis
ticklocsy <- seq(0.4,1,0.1)    # locations of tick marks on y axis
ticklabelsx <- ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("decoding1.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 4, width = 7)

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

# polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$TargetB+CIdecoding$TargetB, meandecoding$TargetB[1201:1]-CIdecoding$TargetB[1201:1]), col=rgb(0,0,0,alpha=0.2), border=NA)

polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$Cue+CIdecoding$Cue, meandecoding$Cue[1201:1]-CIdecoding$Cue[1201:1]), col=flatalpha('blue',0.2), border=NA)

# polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$TargetU+CIdecoding$TargetU, meandecoding$TargetU[1201:1]-CIdecoding$TargetU[1201:1]), col=rgb(1,0,0,alpha=0.2), border=NA)

# polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$CongIncong+CIdecoding$CongIncong, meandecoding$CongIncong[1201:1]-CIdecoding$CongIncong[1201:1]), col=rgb(0,0.5,0,alpha=0.2), border=NA)

# lines(meandecoding$Time,meandecoding$TargetB,col='black')
lines(meandecoding$Time,meandecoding$Cue,col='blue',lwd=3)
# lines(meandecoding$Time,meandecoding$TargetU,col='red')
# lines(meandecoding$Time,meandecoding$CongIncong,col='darkgreen')

i <- which(decodingBFs[,1]>3)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='yellow')
i <- which(decodingBFs[,1]>10)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='orange')
i <- which(decodingBFs[,1]>30)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='red')

legend(-200,1,'Cue side',lwd=3,col='blue',box.lwd=3,bg='white')

dev.off()

postscript("decoding2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 4, width = 7)

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   # create an empty axis of the correct dimensions
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$TargetB+CIdecoding$TargetB, meandecoding$TargetB[1201:1]-CIdecoding$TargetB[1201:1]), col=flatalpha('black',0.2), border=NA)

# polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$Cue+CIdecoding$Cue, meandecoding$Cue[1201:1]-CIdecoding$Cue[1201:1]), col=rgb(0,0,1,alpha=0.2), border=NA)

# polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$TargetU+CIdecoding$TargetU, meandecoding$TargetU[1201:1]-CIdecoding$TargetU[1201:1]), col=rgb(1,0,0,alpha=0.2), border=NA)

polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$CongIncong+CIdecoding$CongIncong, meandecoding$CongIncong[1201:1]-CIdecoding$CongIncong[1201:1]), col=flatalpha('darkgreen',0.2), border=NA)

lines(meandecoding$Time,meandecoding$TargetB,col='black',lwd=3)
# lines(meandecoding$Time,meandecoding$Cue,col='blue')
# lines(meandecoding$Time,meandecoding$TargetU,col='red')
lines(meandecoding$Time,meandecoding$CongIncong,col='darkgreen',lwd=3)

i <- which(decodingBFs[,3]>3)
lines(meandecoding$Time[i],1.08+(0*i),lwd=3,col='yellow')
i <- which(decodingBFs[,3]>10)
lines(meandecoding$Time[i],1.08+(0*i),lwd=3,col='orange')
i <- which(decodingBFs[,3]>30)
lines(meandecoding$Time[i],1.08+(0*i),lwd=3,col='red')

i <- which(decodingBFs[,4]>3)
lines(meandecoding$Time[i],1.02+(0*i),lwd=3,col='yellow')
i <- which(decodingBFs[,4]>10)
lines(meandecoding$Time[i],1.02+(0*i),lwd=3,col='orange')
i <- which(decodingBFs[,4]>30)
lines(meandecoding$Time[i],1.02+(0*i),lwd=3,col='red')

legend(600,1,c('Target side','Target congruency'),lwd=3,col=c('black','darkgreen'),box.lwd=3)

dev.off()

plotlims <- c(-200,1000,-4,4)  
ticklocsx <- seq(-200,1000,200)    # locations of tick marks on x axis
ticklocsy <- seq(-4,4,2)    # locations of tick marks on y axis
ticklabelsx <- ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

postscript("ERP1.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 4.5, width = 8)

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

polygon(c(600,1000,1000,600),c(-4,-4,4,4),border=NA,col='grey90')

lines(c(-200,1000),c(0,0),lty=2)
lines(c(0,0),c(-4,4),lty=3)

for (s in 1:ncol(meancueleft)){
lines(meandecoding$Time,meancueleft[,s],col='black')}

text(1000,3.3,'Cue left',cex=1.5,pos=2)

dev.off()

postscript("ERP2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 4.5, width = 8)

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

polygon(c(600,1000,1000,600),c(-4,-4,4,4),border=NA,col='grey90')

lines(c(-200,1000),c(0,0),lty=2)
lines(c(0,0),c(-4,4),lty=3)

for (s in 1:ncol(meancueright)){
lines(meandecoding$Time,meancueright[,s],col='black')}

text(1000,3.3,'Cue right',cex=1.5,pos=2)

dev.off()

postscript("ERP3.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 4.5, width = 8)

plotlims <- c(-200,1000,-8,8)  
ticklocsy <- seq(-8,8,4)    # locations of tick marks on y axis
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

polygon(c(120,160,160,120),c(-8,-8,8,8),border=NA,col='grey90')

lines(c(-200,1000),c(0,0),lty=2)
lines(c(0,0),c(-8,8),lty=3)

for (s in 1:ncol(meantargetleft)){
lines(meandecoding$Time,meantargetleft[,s],col='black')}

text(1000,7,'Target left',cex=1.5,pos=2)

dev.off()

postscript("ERP4.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 4.5, width = 8)

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)   
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

polygon(c(120,160,160,120),c(-8,-8,8,8),border=NA,col='grey90')

lines(c(-200,1000),c(0,0),lty=2)
lines(c(0,0),c(-8,8),lty=3)

for (s in 1:ncol(meantargetright)){
lines(meandecoding$Time,meantargetright[,s],col='black')}

text(1000,7,'Target right',cex=1.5,pos=2)

dev.off()

xpos <- 1:62
ypos <- 1:62
for (ch in 1:62){
  i <- match(toupper(mnechans[ch]),chandata$Electrode)
  xpos[ch] <- chandata$X_position[i]
  ypos[ch] <- chandata$Y_position[i]
}

ramp2 <- colorRamp(c("darkblue","red"))  # create a ramp from one colour to another
colmatrix2 <- rgb(ramp2(seq(0, 1, length = 101)), max = 255)

## Create data frame to be used for interpolation - the function needs columns labelled x, y, and z
rmax <- 0.55   #specify a maximum boundary for the grid
gridRes <- 100 #specify the interpolation grid resolution

toplot <- colMeans(meancueleft[801:1200,])
toplot[which(is.na(toplot))] <- 0
testDat<- data.frame(x = xpos, y = -ypos, z = toplot)

#Create the interpolation grid
xo <- seq(min(-rmax, testDat$x), max(rmax, testDat$x), length = gridRes)
yo <- seq(max(rmax, testDat$y), min(-rmax, testDat$y), length = gridRes)

interpV4 <- v4Interp(testDat, xo, yo, rmax, gridRes)

zo2 <- as.matrix(interpV4[,2:ncol(interpV4)])

xo2 <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
yo2 <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
outsidecircle <- sqrt(xo2^2 + yo2^2) > 0.51
zo2[outsidecircle] <- 0

tiff("head1.tiff", height = 600, width = 600, units="px", bg="white")

plotlims <- c(-rmax,rmax,-rmax,rmax) 
par(pty="s")  # make axis square
plot(x=NULL,y=NULL,ann=FALSE, axes=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
image(xo,xo,zo2,col=colmatrix2,zlim=c(-3,3),add=TRUE,useRaster=TRUE)
maskx <- c(chandata$OutlineX[1:51]*2.2,chandata$OutlineX[51:1])
masky <- c(chandata$OutlineY[1:51]*2.2,chandata$OutlineY[51:1])
polygon(maskx,masky,border=NA,col="white")
maskx <- c(chandata$OutlineX[51:101]*2.2,chandata$OutlineX[101:51])
masky <- c(chandata$OutlineY[51:101]*2.2,chandata$OutlineY[101:51])
polygon(maskx,masky,border=NA,col="white")

# blackelectrodes <- match(toupper(targetelectrodes),toupper(as.character(electrodes[3:66])))
# points(xpos[blackelectrodes],ypos[blackelectrodes],pch=16,col="black",cex=2)

lines(chandata$OutlineX,chandata$OutlineY,col="black",lwd=2)
lines(chandata$NoseX,chandata$NoseY,col="black",lwd=2)
lines(chandata$LearX,chandata$LearY,col="black",lwd=2)
lines(chandata$RearX,chandata$RearY,col="black",lwd=2)

dev.off()

toplot <- colMeans(meancueright[801:1200,])
toplot[which(is.na(toplot))] <- 0
testDat<- data.frame(x = xpos, y = -ypos, z = toplot)

#Create the interpolation grid
xo <- seq(min(-rmax, testDat$x), max(rmax, testDat$x), length = gridRes)
yo <- seq(max(rmax, testDat$y), min(-rmax, testDat$y), length = gridRes)

interpV4 <- v4Interp(testDat, xo, yo, rmax, gridRes)

zo2 <- as.matrix(interpV4[,2:ncol(interpV4)])

xo2 <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
yo2 <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
outsidecircle <- sqrt(xo2^2 + yo2^2) > 0.51
zo2[outsidecircle] <- 0

tiff("head2.tiff", height = 600, width = 600, units="px", bg="white")

plotlims <- c(-rmax,rmax,-rmax,rmax) 
par(pty="s")  # make axis square
plot(x=NULL,y=NULL,ann=FALSE, axes=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
image(xo,xo,zo2,col=colmatrix2,zlim=c(-3,3),add=TRUE,useRaster=TRUE)
maskx <- c(chandata$OutlineX[1:51]*2.2,chandata$OutlineX[51:1])
masky <- c(chandata$OutlineY[1:51]*2.2,chandata$OutlineY[51:1])
polygon(maskx,masky,border=NA,col="white")
maskx <- c(chandata$OutlineX[51:101]*2.2,chandata$OutlineX[101:51])
masky <- c(chandata$OutlineY[51:101]*2.2,chandata$OutlineY[101:51])
polygon(maskx,masky,border=NA,col="white")

# blackelectrodes <- match(toupper(targetelectrodes),toupper(as.character(electrodes[3:66])))
# points(xpos[blackelectrodes],ypos[blackelectrodes],pch=16,col="black",cex=2)

lines(chandata$OutlineX,chandata$OutlineY,col="black",lwd=2)
lines(chandata$NoseX,chandata$NoseY,col="black",lwd=2)
lines(chandata$LearX,chandata$LearY,col="black",lwd=2)
lines(chandata$RearX,chandata$RearY,col="black",lwd=2)

dev.off()

toplot <- colMeans(meantargetleft[321:360,])
toplot[which(is.na(toplot))] <- 0
testDat<- data.frame(x = xpos, y = -ypos, z = toplot)

#Create the interpolation grid
xo <- seq(min(-rmax, testDat$x), max(rmax, testDat$x), length = gridRes)
yo <- seq(max(rmax, testDat$y), min(-rmax, testDat$y), length = gridRes)

interpV4 <- v4Interp(testDat, xo, yo, rmax, gridRes)

zo2 <- as.matrix(interpV4[,2:ncol(interpV4)])

xo2 <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
yo2 <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
outsidecircle <- sqrt(xo2^2 + yo2^2) > 0.51
zo2[outsidecircle] <- 0

zo2[which(zo2 > 3)] <- 3
zo2[which(zo2 < -3)] <- -3

tiff("head3.tiff", height = 600, width = 600, units="px", bg="white")

plotlims <- c(-rmax,rmax,-rmax,rmax) 
par(pty="s")  # make axis square
plot(x=NULL,y=NULL,ann=FALSE, axes=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
image(xo,xo,zo2,col=colmatrix2,zlim=c(-3,3),add=TRUE,useRaster=TRUE)
maskx <- c(chandata$OutlineX[1:51]*2.2,chandata$OutlineX[51:1])
masky <- c(chandata$OutlineY[1:51]*2.2,chandata$OutlineY[51:1])
polygon(maskx,masky,border=NA,col="white")
maskx <- c(chandata$OutlineX[51:101]*2.2,chandata$OutlineX[101:51])
masky <- c(chandata$OutlineY[51:101]*2.2,chandata$OutlineY[101:51])
polygon(maskx,masky,border=NA,col="white")

# blackelectrodes <- match(toupper(targetelectrodes),toupper(as.character(electrodes[3:66])))
# points(xpos[blackelectrodes],ypos[blackelectrodes],pch=16,col="black",cex=2)

lines(chandata$OutlineX,chandata$OutlineY,col="black",lwd=2)
lines(chandata$NoseX,chandata$NoseY,col="black",lwd=2)
lines(chandata$LearX,chandata$LearY,col="black",lwd=2)
lines(chandata$RearX,chandata$RearY,col="black",lwd=2)

dev.off()

toplot <- colMeans(meantargetright[321:360,])
toplot[which(is.na(toplot))] <- 0
testDat<- data.frame(x = xpos, y = -ypos, z = toplot)

#Create the interpolation grid
xo <- seq(min(-rmax, testDat$x), max(rmax, testDat$x), length = gridRes)
yo <- seq(max(rmax, testDat$y), min(-rmax, testDat$y), length = gridRes)

interpV4 <- v4Interp(testDat, xo, yo, rmax, gridRes)

zo2 <- as.matrix(interpV4[,2:ncol(interpV4)])

xo2 <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
yo2 <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
outsidecircle <- sqrt(xo2^2 + yo2^2) > 0.51
zo2[outsidecircle] <- 0

zo2[which(zo2 > 3)] <- 3
zo2[which(zo2 < -3)] <- -3

tiff("head4.tiff", height = 600, width = 600, units="px", bg="white")

plotlims <- c(-rmax,rmax,-rmax,rmax) 
par(pty="s")  # make axis square
plot(x=NULL,y=NULL,ann=FALSE, axes=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])  
image(xo,xo,zo2,col=colmatrix2,zlim=c(-3,3),add=TRUE,useRaster=TRUE)
maskx <- c(chandata$OutlineX[1:51]*2.2,chandata$OutlineX[51:1])
masky <- c(chandata$OutlineY[1:51]*2.2,chandata$OutlineY[51:1])
polygon(maskx,masky,border=NA,col="white")
maskx <- c(chandata$OutlineX[51:101]*2.2,chandata$OutlineX[101:51])
masky <- c(chandata$OutlineY[51:101]*2.2,chandata$OutlineY[101:51])
polygon(maskx,masky,border=NA,col="white")

# blackelectrodes <- match(toupper(targetelectrodes),toupper(as.character(electrodes[3:66])))
# points(xpos[blackelectrodes],ypos[blackelectrodes],pch=16,col="black",cex=2)

lines(chandata$OutlineX,chandata$OutlineY,col="black",lwd=2)
lines(chandata$NoseX,chandata$NoseY,col="black",lwd=2)
lines(chandata$LearX,chandata$LearY,col="black",lwd=2)
lines(chandata$RearX,chandata$RearY,col="black",lwd=2)

dev.off()

  PostScriptTrace('decoding1.ps')
  p1 <- readPicture('decoding1.ps.xml')
  PostScriptTrace('decoding2.ps')
  p2 <- readPicture('decoding2.ps.xml')
  PostScriptTrace('ERP1.ps')
  p3 <- readPicture('ERP1.ps.xml')
  PostScriptTrace('ERP2.ps')
  p4 <- readPicture('ERP2.ps.xml')
  PostScriptTrace('ERP3.ps')
  p5 <- readPicture('ERP3.ps.xml')
  PostScriptTrace('ERP4.ps')
  p6 <- readPicture('ERP4.ps.xml')

  h1 <- readTIFF('head1.tiff')
  h2 <- readTIFF('head2.tiff')
  h3 <- readTIFF('head3.tiff')
  h4 <- readTIFF('head4.tiff')

    pdf(paste0(figdir,"Exp1summary.pdf"), bg="transparent", height = 10, width = 15)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  # insert the head plots first so the white border doesn't overlap the other graphs
  aspratio <- 10/15  # this is the aspect ratio of the output pdf
  imwidth <- 0.25
  xstart <- 0.34
  ystart <- 0.34
  rasterImage(h1,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth) 
  ystart <- 0.01
  rasterImage(h2,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth) 
  xstart <- 0.87
  ystart <- 0.34
  rasterImage(h3,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth) 
  ystart <- 0.01
  rasterImage(h4,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)

    grid.picture(p1,x=0.25,y=0.78,width=0.5,height=1)
    grid.picture(p2,x=0.75,y=0.78,width=0.5,height=1)
    grid.picture(p3,x=0.2,y=0.45,width=0.35,height=1)
    grid.picture(p4,x=0.2,y=0.15,width=0.35,height=1)
    grid.picture(p5,x=0.7,y=0.45,width=0.35,height=1)
    grid.picture(p6,x=0.7,y=0.15,width=0.35,height=1)

    text(0.0,1,'(a)',cex=3,adj=0.5)
    text(0.53,1,'(d)',cex=3,adj=0.5)
    text(0.0,0.6,'(b)',cex=3,adj=0.5)
    text(0.53,0.6,'(e)',cex=3,adj=0.5)
    text(0.0,0.28,'(c)',cex=3,adj=0.5)
    text(0.53,0.28,'(f)',cex=3,adj=0.5)

      dev.off()

  file.remove(c('decoding1.ps','decoding2.ps','ERP1.ps','ERP2.ps','ERP3.ps','ERP4.ps','head1.tiff','head2.tiff','head3.tiff','head4.tiff'))

  file.remove(c('decoding1.ps.xml','decoding2.ps.xml','ERP1.ps.xml','ERP2.ps.xml','ERP3.ps.xml','ERP4.ps.xml'))

}

```

```{r makeplotExpt2EEG}
#| include: false
#| output: false
#| warning: false

if (processdata>0){

load('local/Exp2GroupStim.RData')
  stimdecoding <- meandecoding
  stimCI <- CIdecoding
  stimBFs <- decodingBFs2
load('local/Exp2GroupNoStim.RData')
  nostimdecoding <- meandecoding
  nostimCI <- CIdecoding  
  nostimBFs <- decodingBFs2

pdf("Figures/Exp2summary.pdf", height = 8, width = 12)

par(mfrow=c(3,2))

load('local/Exp2Group.RData')

plotlims <- c(-200,1000,0.4,1.1)  
ticklocsx <- seq(-200,1000,200)    # locations of tick marks on x axis
ticklocsy <- seq(0.4,1,0.1)    # locations of tick marks on y axis
ticklabelsx <- ticklocsx        # set labels for x ticks
ticklabelsy <- ticklocsy    # set labels for y ticks

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.5)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main='(a) Cue side (all participants)',cex.main=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$Cue+CIdecoding$Cue, meandecoding$Cue[1201:1]-CIdecoding$Cue[1201:1]), col=addalpha('blue',0.2), border=NA)

lines(meandecoding$Time,meandecoding$Cue,col='blue',lwd=3)

i <- which(decodingBFs2[,1]>3)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='yellow')
i <- which(decodingBFs2[,1]>10)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='orange')
i <- which(decodingBFs2[,1]>30)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='red')

legend(-200,1,'Cue side',cex=1.5,lwd=3,col='blue',box.lwd=2,bg='white')

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.5)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main='(b) Target side (all participants)',cex.main=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$TargetB+CIdecoding$TargetB, meandecoding$TargetB[1201:1]-CIdecoding$TargetB[1201:1]), col=addalpha('black',0.2), border=NA)

polygon(meandecoding$Time[c(1:1201,1201:1)],c(meandecoding$CongIncong+CIdecoding$CongIncong, meandecoding$CongIncong[1201:1]-CIdecoding$CongIncong[1201:1]), col=addalpha('darkgreen',0.2), border=NA)

lines(meandecoding$Time,meandecoding$TargetB,col='black',lwd=3)
lines(meandecoding$Time,meandecoding$CongIncong,col='darkgreen',lwd=3)

i <- which(decodingBFs2[,4]>3)
lines(meandecoding$Time[i],1.08+(0*i),lwd=3,col='yellow')
i <- which(decodingBFs2[,4]>10)
lines(meandecoding$Time[i],1.08+(0*i),lwd=3,col='orange')
i <- which(decodingBFs2[,4]>30)
lines(meandecoding$Time[i],1.08+(0*i),lwd=3,col='red')

i <- which(decodingBFs2[,3]>10)
lines(meandecoding$Time[i],1.02+(0*i),lwd=3,col='yellow')
i <- which(decodingBFs2[,3]>30)
lines(meandecoding$Time[i],1.02+(0*i),lwd=3,col='orange')
i <- which(decodingBFs2[,3]>100)
lines(meandecoding$Time[i],1.02+(0*i),lwd=3,col='red')

legend(500,1,c('Target side','Target congruency'),lwd=3,col=c('black','darkgreen'),cex=1.5,box.lwd=2)

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.5)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main='(c) Cue side (group comparison)',cex.main=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

polygon(stimdecoding$Time[c(1:1201,1201:1)],c(stimdecoding$Cue+stimCI$Cue, stimdecoding$Cue[1201:1]-stimCI$Cue[1201:1]), col=addalpha('saddlebrown',0.2), border=NA)
polygon(nostimdecoding$Time[c(1:1201,1201:1)],c(nostimdecoding$Cue+nostimCI$Cue, nostimdecoding$Cue[1201:1]-nostimCI$Cue[1201:1]), col=addalpha('blueviolet',0.2), border=NA)

lines(stimdecoding$Time,stimdecoding$Cue,col='saddlebrown',lwd=3)
lines(nostimdecoding$Time,nostimdecoding$Cue,col='blueviolet',lwd=3)

i <- which(groupcomp[,1]>3)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='yellow')
i <- which(groupcomp[,1]>10)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='orange')
i <- which(groupcomp[,1]>30)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='red')

legend(-200,1,c('Stimming group','No stimming group'),lwd=3,col=c('saddlebrown','blueviolet'),cex=1.5,box.lwd=2,bg='white')

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.5)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main='(d) Cue congruency (group comparison)',cex.main=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

polygon(stimdecoding$Time[c(1:1201,1201:1)],c(stimdecoding$CongIncong+stimCI$CongIncong, stimdecoding$CongIncong[1201:1]-stimCI$CongIncong[1201:1]), col=addalpha('saddlebrown',0.2), border=NA)
polygon(nostimdecoding$Time[c(1:1201,1201:1)],c(nostimdecoding$CongIncong+nostimCI$CongIncong, nostimdecoding$CongIncong[1201:1]-nostimCI$CongIncong[1201:1]), col=addalpha('blueviolet',0.2), border=NA)

lines(stimdecoding$Time,stimdecoding$CongIncong,col='saddlebrown',lwd=3)
lines(nostimdecoding$Time,nostimdecoding$CongIncong,col='blueviolet',lwd=3)

i <- which(groupcomp[,2]>3)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='yellow')
i <- which(groupcomp[,2]>10)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='orange')
i <- which(groupcomp[,2]>30)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='red')

legend(-200,1,c('Stimming group','No stimming group'),lwd=3,col=c('saddlebrown','blueviolet'),cex=1.5,box.lwd=2,bg='white')

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.5)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main='(e) Cue side (stimming group)',cex.main=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

polygon(stimdecoding$Time[c(1:1201,1201:1)],c(stimdecoding$CueNostim+stimCI$CueNostim, stimdecoding$CueNostim[1201:1]-stimCI$CueNostim[1201:1]), col=addalpha('goldenrod',0.2), border=NA)
polygon(stimdecoding$Time[c(1:1201,1201:1)],c(stimdecoding$CueStim+stimCI$CueStim, stimdecoding$CueStim[1201:1]-stimCI$CueStim[1201:1]), col=addalpha('forestgreen',0.2), border=NA)

lines(nostimdecoding$Time,stimdecoding$CueNostim,col='goldenrod',lwd=3)
lines(nostimdecoding$Time,stimdecoding$CueStim,col='forestgreen',lwd=3)

i <- which(stimBFs[,7]>3)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='yellow')
i <- which(stimBFs[,7]>10)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='orange')
i <- which(stimBFs[,7]>30)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='red')

legend(-200,1,c('No stimming condition','Stimming condition'),lwd=3,col=c('goldenrod','forestgreen'),cex=1.5,box.lwd=2,bg='white')

plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])   
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklabelsx, side = 1, at=ticklocsx, line=0.5)     # add the tick labels
mtext(text = ticklabelsy, side = 2, at=ticklocsy, line=0.2, las=1)  
title(xlab="Time (ms)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)   
title(ylab="Proportion correct", col.lab=rgb(0,0,0), line=2, cex.lab=1.5)
title(main='(f) Cue side (non-stimming group)',cex.main=1.5)

lines(c(-200,1000),c(0.5,0.5),lty=2)
lines(c(0,0),c(0,1),lty=3)

polygon(nostimdecoding$Time[c(1:1201,1201:1)],c(nostimdecoding$CueNostim+nostimCI$CueNostim, nostimdecoding$CueNostim[1201:1]-nostimCI$CueNostim[1201:1]), col=addalpha('goldenrod',0.2), border=NA)
polygon(nostimdecoding$Time[c(1:1201,1201:1)],c(nostimdecoding$CueStim+nostimCI$CueStim, nostimdecoding$CueStim[1201:1]-nostimCI$CueStim[1201:1]), col=addalpha('forestgreen',0.2), border=NA)

lines(nostimdecoding$Time,nostimdecoding$CueNostim,col='goldenrod',lwd=3)
lines(nostimdecoding$Time,nostimdecoding$CueStim,col='forestgreen',lwd=3)

i <- which(nostimBFs[,7]>3)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='yellow')
i <- which(nostimBFs[,7]>10)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='orange')
i <- which(nostimBFs[,7]>30)
lines(meandecoding$Time[i],1.05+(0*i),lwd=3,col='red')

legend(-200,1,c('No stimming condition','Stimming condition'),lwd=3,col=c('goldenrod','forestgreen'),cex=1.5,box.lwd=2,bg='white')

dev.off()
}

```

```{r dostatsExpt1EEG}
#| include: false
#| output: false
#| warning: false

load('local/Exp1Group.RData')

maxcue <- round(max(meandecoding$Cue)*100)
i <- which(meandecoding$Cue==max(meandecoding$Cue))
maxcuetime <- meandecoding$Time[i]
maxtarget <- round(max(meandecoding$TargetB)*100)
maxcong <- round(max(meandecoding$CongIncong)*100)

load('local/Expt2resps.RData')

meanRTnostimming <- round(10^(mean(allRT[,1:2])/20))
meanRTstimming <- round(10^(mean(allRT[,3:4])/20))

```

# Abstract

# Introduction

The ability to orient and control the deployment of our attention to spatial locations or features of our environment is essential to our ability to perform tasks, focusing on relevant stimuli and disregarding those that are unnecessary for the task at hand [@Knudsen2007; @Hopfinger2000; @Rueda2011]. A classic technique for studying the deployment of attention in space is the Posner cueing paradigm, which demonstrates that it is possible for attention to be deployed to spatial locations independently of eye movements [@Posner1980].

Individuals with neurodevelopmental disorders such as attention deficit hyperactivity disorder (ADHD) and autism present with deficits in attention (American Psychiatric Association, 2013; [@Allen2001]. It has been suggested that repetitive stereotypies, or ‘stimming’ behaviours, are linked to these deficits in attention -- in the few studies where they have been asked, individuals who stim report that it benefits their ability to focus, and that the behaviours aid in emotional regulation [@Kapp2019; @Steward2015]. It has been proposed that an improved ability to focus through stimming may have a neural basis [@McCarty2021]. Building on these theories and research, we tested the effect of stimming on performance in a Posner cueing task.

The Posner cueing paradigm [@Posner1980] involves the participant attending to a central fixation point, then receiving a cue (which may be exogenous, where attention is captured through bottom-up processes; or endogenous, where attention is controlled by top-down processes, such as attending selectively to cues of the correct colour [@Leblanc2008]) indicating that a stimulus is about to appear either to the left or the right of the fixation point [@Thiery2016]. The assumption underlying the task is that if participants successfully orient their attention to the cued location, then their reaction times will be shorter and responses more accurate for congruent trials (where the stimulus appears at the cued location) than for incongruent trials (where the stimulus appears at the uncued location); this has been demonstrated extensively [@Thiery2016]. The reaction time difference was found by @Posner1978 to be 25ms faster for congruent trials and 40ms slower for incongruent trials, compared to a neutral cue condition.

This paradigm has been combined with EEG to investigate the neural basis of spatial attention. For example, @Landau2007 investigated gamma-band activity during a Posner cueing task, comparing congruent and incongruent trials. Gamma range responses were associated with voluntary shifts in attention, e.g., attending to the left of the fixation point following a left-pointing arrow cue, but were absent when attention was involuntarily captured by the appearance of a stimulus in the uncued location on incongruent trials; this suggests that voluntary and involuntary processes of attention deployment are controlled by separate mechanisms. @Thiery2016 investigated two specific ERPs, the N2pc component (a negative response typically seen at around 200ms post stimulus, in the posterior-contralateral region) and the SPCN component (sustained posterior contralateral negativity). They used multivariate pattern analysis and used the decoder accuracy to determine that these two components were able to predict the focus of participants’ spatial attention. This was modulated by the varied distance of the stimulus from the central fixation (further being better decoded).

A combined EEG and TMS study by @Capotosto2012 demonstrated that the application of TMS to the right intraparietal sulcus (IPS) impairs target detection, particularly on incongruent trials. They additionally noted that the P3 response amplitude was affected by this; on incongruent trials its amplitude was significantly lower, and on congruent trials it was significantly higher. The IPS has been previously identified as an important region in attentional deployment by @Vossel2009 in an fMRI study identifying regions that are selectively active on congruent and incongruent trials. The right IPS and, additionally, the right inferior frontal gyrus were implicated in attentional processes related to both types of trial. @Peelen2004 concluded that both endogenous and exogenous orienting cues involve the same network in their fMRI study, implicating frontal and parietal regions (the premotor cortex, posterior parietal cortex, medial frontal cortex, and right inferior frontal cortex). @Fitzgerald2015 conducted an MRI study using the Posner cueing task and found that activation in the ventral attention network differed between individuals with autism and controls; individuals with autism also showed weaker functional connectivity in the dorsal attentional network compared to their neurotypical counterparts.

A number of developmental disorders involve deficits in attention, notably ADHD and autism. In the case of ADHD, attention deficits are a defining feature of the condition (American Psychological Association, 2013). For autism, attention deficits are not an explicit part of the diagnostic criteria (although restricted interests, which are included, are somewhat related), but both attentional deficits and strengths are widely reported in autistic individuals. They are typically more distractible, demonstrating ‘underselective’ attention, but also demonstrate ‘overselective’ attention where they attend intensely to a more limited subset of stimuli. This ‘overselective’ attention has been suggested to be linked to restricted interests, which is also sometimes referred to as ‘hyperfocus’, where a task is the subject of intense focus for the individual [e.g., @Dupuis2022]. These differences appear to have a neural basis -- they have been linked to unusually wide sulci in the parietal lobe [@Allen2001]. Hyperfocus is also widely self-reported in ADHD, despite a lack of clinical research into the phenomenon [@Hupfeld2019].

These marked differences in attentional deployment seen in both conditions are worth considering in the framework of the transdiagnostic approach, which seeks to identify core deficiencies that result in the presentation of traits from multiple disorders, instead of approaching different disorders as entirely separate even within the same individual [@Newby2015]. ADHD is the most common comorbidity of autism, and even without a dual diagnosis, 30-80% of individuals with autism present symptoms of ADHD, and 20-60% of children with ADHD exhibit autism-like traits [@Zhang2022]. Some shared deficiencies that have already been identified in the case of autism and ADHD include fine motor function and verbal fluency, alongside some structural differences in the brain [@Zhang2022].

Stimming behaviours are commonly seen in individuals with diagnoses of autism and/or ADHD, and involve repetitive movements, typically featuring the arms, hands, or entire body [@Chadehumbe2018], for example hand flapping [@Kapp2019]. They are also observed in neurotypical individuals, but at much lower rates than in neurodiverse individuals [@Chadehumbe2018]. The behaviours are typically defined as involuntary, but pleasant [@Chadehumbe2018; @Kapp2019; @Steward2015], and are associated with emotional states such as stress, boredom, concentration, or excitement [@Mackenzie2018]. They are often defined as purposeless [e.g., @Tan1997] but self-reports from individuals with autism claim that stimming aids the regulation of their emotions [@Kapp2019; @Steward2015].

A type of vocal stimming called echolalia has been consistently demonstrated to have benefits for autistic children in the acquisition of language [@Charlop1983; @Pruccoli2021]. This is one of the few instances that stimming has been researched in terms of its potential function. Echolalic utterances, which typically involve repeating words spoken by another, have also been identified as having predominantly communicative functions [@Prizant1981]. This opens up an avenue to research potential benefits that other types of stimming may provide.

Current research into stimming is largely focussed on designing interventions to eliminate the behaviours, for example the use of weighted vests [@FertelDaly2001]. Taking this approach without first attempting to understand if there are benefits to the behaviour seems ill-advised. @McCarty2021, for example, proposed based on self-reports of the benefits of stimming (focus, coping with overwhelming sensory input, and relaxation) that regular motor movements may generate (or be a by-product of) rhythmic oscillations in the motor cortex, which entrain oscillations in the sensory cortex. This entrainment might improve information transfer in the sensory cortex, effectively normalising the atypical motor and sensory oscillations observed in the brains of autistic individuals both during tasks [@Milne2009; @Snijders2013; @Murphy2017] and at rest [@Wang2013; @Berman2015].

There is heavy stigma around stimming behaviours, with many individuals reporting they have been explicitly instructed not to stim, despite attempts at repression causing them some discomfort [@Steward2015]. Individuals with autism state that they believe stimming should not be stigmatised in this way, and that they should be allowed to act in the way that feels natural to them [@Kapp2019]. Research into potential benefits of stimming is important as a step towards dispelling this stigma, and discouraging interventions that aim to eliminate the behaviour.

This study involved two experiments. In the first experiment, participants completed a Posner cueing task while EEG recordings were taken; a pattern classifier then attempted to differentiate, using the participants’ brain activity, whether they attended to the left vs. to the right following the cue onset. This is similar to the methodology used by @Thiery2016\: by measuring the timecourse of classifier accuracy, we are provided with an index of the deployment of spatial attention.

The second study extended this paradigm by adding a stimming condition. The accuracy of the classifier was compared between conditions where the participants were permitted to stim during the trials, or instructed to keep still. In this case, we used decoding accuracy to assess whether stimming provided a benefit in terms of the deployment of spatial attention; if there was a benefit, the classifier accuracy would be higher, as the EEG signal would differ more between conditions due to the greater focus of attention in different spatial locations. For the purposes of our analyses, due to an insufficiently sized and heterogeneous clinical group, we grouped the participants based on whether or not they reported stimming in their everyday lives; as discussed previously, neurotypicals often engage in stimming [@Chadehumbe2018], and 64% of our participants reported in the pre-experiment questionnaire that they typically engaged in stimming.

# Methods

## Participants

41 participants took part in Experiment 1, but four produced very noisy data so were excluded, leaving a sample of 37 participants (20 male, 17 female). A total of 22 participants took part in Experiment 2 (8 male, 14 female). Participants in Experiment 2 were asked to report diagnoses of neurodevelopmental disorders – 17 had no diagnosis, 2 had a diagnosis of autism, and 3 had a diagnosis of ADHD. 15/22 participants further reported they typically engaged in stimming behaviours, while 7 reported that they did not.

## Apparatus and stimuli

Stimuli were displayed on a ViewPixx 3D display (VPixx Technologies Inc., Quebec, Canada) with a diagonal extent of 24 inches, a refresh rate of 120Hz, and a resolution of 1920x1080 pixels. The screen was viewed by the participant from a distance of 57cm, resulting in a spatial resolution of 36 pixels per degree of visual angle. The display was gamma corrected using a Minolta LS110 photometer to ensure a linear luminance response. Stimuli were generated and displayed using a Mac Pro computer running Matlab, with the experiment script making use of functions from Psychtoolbox 3 [@Kleiner2007; @Brainard1997; @Pelli1997].

In both experiments, EEG recordings were acquired using an ANT Neuroscan 64-channel EEG system (ANT Neuro). Electrodes were mounted in a waveguard cap, arranged according to the 10/20 system, and were referenced to the whole head average with the ground electrode located at position *AFz*. EEG signals were recorded at 1kHz. In Experiment 2, participants’ eye movements were also recorded using an EyeLink 1000 (SR Research Ltd., Ottawa, Canada) eye tracker, also sampling at 1kHz. Low-latency digital triggers were sent from the stimulus computer to the EEG system using an 8-bit parallel cable.

Target stimuli in both experiments were horizontal sine-wave gratings with a diameter of 4 degrees, and a spatial frequency of 2c/deg. The target appeared offset by 8 degrees to either the left or the right of the central fixation point. The base contrast of the target was 50\%, and either the upper or lower half of the grating contained a contrast increment of a further 10\%. Participants were asked to indicate whether the upper or lower half of the grating contained the increment using arrow keys on the computer keyboard. The fixation cross was 0.25 $\times$ 0.25 degrees, and was displayed throughout.

Each trial was preceded by a cue. In Experiment 1 the cue was a face with eyes pointing left or right. In Experiment 2 the cue was an arrow pointing left or right. The face cues used in Experiment 1 were the average of 22 female faces taken from the NimStim database [@Tottenham2009] and were digitally altered so that the eyes were directed to the left or the right. The face had a width of 3 degrees and a height of 4 degrees and was displayed in the centre of the screen. The arrow cue used in Experiment 2 was 1.4 degrees long and was displayed in the centre of the screen.

## Procedure

Participants in both experiments completed four blocks of the Posner cueing task, with each block consisting of 200 trials. They were instructed to keep their eyes focused on the central fixation cross, and to attend to the gratings without shifting their gaze. During each trial, participants viewed a fixation cross, which remained on the screen throughout the trial (excluding times when the cue appeared on the screen). The cue stimulus appeared for 200ms, followed by a randomly determined gap of 400-600ms before the grating appeared on the screen. The target grating was then displayed for 200ms. Participants were required to report whether the upper or lower half of the grating appeared higher in contrast; following their response, a randomly determined interval drawn from a normal distribution (average 1000ms, with a standard deviation of ±200ms) preceded the next appearance of the cue. This procedure is illustrated in @fig-methodsfig for the face cue (panel a) and the arrow cue (panel b). Cue congruence was 75\% in both experiments, meaning that each participant contributed a total of 600 congruent trials and 200 incongruent trials across the experiment.

```{r methodsfig}
#| include: true
#| output: true
#| label: fig-methodsfig
#| out.width: '95%'
#| fig.cap: 'Illustration of trial procedure for both experiments. In panel (a), the cue was a face with eyes pointing to either the left (as pictured) or the right. In panel (b), the cue was an arrow. Target gratings had a contrast of 50\%, plus a 10\% increment in either the upper or lower half. Note that images are not drawn to scale in this schematic to aid visibility - see text for actual dimensions.'

knitr::include_graphics(paste0(figdir,'MethodsFigure.pdf'))

```

In Experiment 2, participants were additionally instructed to stim, using a movement restricted to one hand only, in half of the blocks. This was to ensure minimal distortion of the EEG signal resulting from movement. The order in which participants completed stimming or non-stimming blocks was counterbalanced; half the participants were instructed to stim on blocks 1 and 3, while the other half were instructed to stim on blocks 2 and 4.

## Data analysis

Raw EEG data were converted into EEGlab .set format using Matlab and the EEGlab toolbox [@Delorme2004]. We then used MNE python for the main analysis. Data were bandpass filtered between 0.2 and 30Hz, and then epoched using the event timestamps. We performed classification using a linear classifier with 10-fold cross-validation on the voltage data across all electrodes (excluding electrodes *M1* and *M2*), independently at each time point and for each participant. We then averaged classification accuracy across participants to generate the timecourses used in the main analysis. We calculate Bayes factors [@Rouder2009] to compare classification accuracy to chance performance (50\% correct) and to make between conditions and groups, but also report frequentist tests for comparison. We express Bayes factors in logarithmic units to account for extreme values

## Open science practices

Raw and processed data are available for both experiments at the project repository on the Open Science Framework website: http://doi.org/10.17605/OSF.IO/YUHZ4. Experiment code and analysis code are provided in the linked Github repository (https://github.com/bakerdh/attentionstimming), which contains a fully computationally reproducible document written in Quarto, that generates this manuscript using the raw data.

# Results

## Experiment 1

We first analysed the behavioural responses during the experiment, comparing accuracy and reaction times between stimuli shown in the cued (i.e. a congruent target) and uncued (an incongruent target) location. There was a highly significant effect of cue congruency on both accuracy ($log_{10}BF_{10}$ = `r round(log10(extractBF(ACCBF)$bf),digits=2)`, *t* = `r round(ACCtest$statistic,digits=2)`, *p* `r p_format(ACCtest$p.value,digits=3,accuracy=0.001)`, *d* = `r round(ACCd,digits=2)`) and reaction time ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTBF)$bf),digits=2)`, *t* = `r round(RTtest$statistic,digits=2)`, *p* `r p_format(RTtest$p.value,digits=3,accuracy=0.001)`, *d* = `r round(RTd,digits=2)`), as shown in @fig-Expt1resps. Responses to targets in the cued location were both faster ($M_{diff}$ = `r round(RTdiff)` ms) and more accurate ($M_{diff}$ = `r round(accdiff)`\%) than for those in the uncued location, showing a classic Posner cueing effect.

```{r plotExpt1resps}
#| include: true
#| output: true
#| label: fig-Expt1resps
#| out.width: '60%'
#| fig.cap: 'Summary of reaction time and accuracy data for the cueing task in Experiment 1. Small points show data for individual participants (N = 39), and larger points give the group means. Error bars indicate 95% confidence intervals.'

knitr::include_graphics(paste0(figdir,'Exp1resps.pdf'))

```

We next explored the EEG data in response to both the cue presentation and the target presentation. There were typical evoked potentials in response to both types of stimulus, as summarised in [@fig-Expt1EEG]b,c,e,f. Patterns of voltage across the scalp showed similar lateralisation effects both for cues pointing to either the left or right, and for targets presented to either the left or right. Multivariate pattern analysis was able to decode the side of cue presentation from around 300ms after cue onset (see [@fig-Expt1EEG]a), with a maximum accuracy of `r maxcue`%, and a plateau of high performance from around 500-1000ms. This relatively late onset of above-chance classification likely reflects the deployment of covert spatial attention to one side of the visual field, and is therefore a neural index of attentional focus. The pattern classifier could decode the side of target presentation with much greater accuracy (maximum of `r maxtarget`%), from around 100ms after target onset (black curve in [@fig-Expt1EEG]d). Finally, we were able to decode cue congruency (i.e. whether the target presented was on the side indicated by the cue) with a maximum accuracy of `r maxcong`%, and a similar timecourse to the decoding of target side (green curve in [@fig-Expt1EEG]d). Overall, these results indicate that our EEG data are sufficiently sensitive to provide an index of the deployment of spatial attention.

```{r plotExpt1EEG}
#| include: true
#| output: true
#| label: fig-Expt1EEG
#| fig.cap: 'Summary of EEG data from Experiment 1. Panel (a) shows decoding accuracy of a pattern classification algorithm trained on responses to the cue stimulus (Time 0 is the cue onset). ERPs and scalp topographies are shown for cues pointing left (b) and right (c), with the shaded regions indicating the time period over which voltages were averaged to generate the scalp plots. Panel (d) shows decoding accuracy for classifying target location (black curve) or cue congruency (green curve) in response to the target (Time 0 is the target onset). ERPs and scalp topographies are shown for targets on the left (e) and right (f) side of fixation, with the shaded regions indicating the time period over which voltages were averaged to generate the scalp plots. Shaded regions in panels (a,d) indicate 95% confidence intervals.'

knitr::include_graphics(paste0(figdir,'Exp1summary.pdf'))

```

## Experiment 2

In Experiment 2, we replicated the basic design of Experiment 1, but this time with an added 'stimming' condition, and with participants divided into groups based on their self-reported stimming behaviour. Pooling across all participants, for accuracy there was no effect of cue congruency ($log_{10}BF_{10}$ = `r round(log10(extractBF(accAOVBF)$bf[2]),digits=2)`, *F* = `r round(accAOVresults$ANOVA[2,4],digits=2)`, *p* = `r p_format(accAOVresults$ANOVA[2,5],digits=3,accuracy=0.001)`, $\eta^2_G$ = `r round(accAOVresults$ANOVA[2,7],digits=2)`), stimming condition ($log_{10}BF_{10}$ = `r round(log10(extractBF(accAOVBF)$bf[1]),digits=2)`, *F* = `r round(accAOVresults$ANOVA[1,4],digits=2)`, *p* = `r p_format(accAOVresults$ANOVA[1,5],digits=3,accuracy=0.001)`, $\eta^2_G$ = `r round(accAOVresults$ANOVA[1,7],digits=2)`), nor an interaction between the two ($log_{10}BF_{10}$ = `r round(log10(extractBF(accAOVBF)$bf[3]),digits=2)`, *F* = `r round(accAOVresults$ANOVA[3,4],digits=2)`, *p* = `r p_format(accAOVresults$ANOVA[3,5],digits=3,accuracy=0.001)`, $\eta^2_G$ = `r round(accAOVresults$ANOVA[3,7],digits=2)`). For reaction time, there were significant effects of cue congruency ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTAOVBF)$bf[2]),digits=2)`, *F* = `r round(RTAOVresults$ANOVA[2,4],digits=2)`, *p* `r p_format(RTAOVresults$ANOVA[2,5],digits=3,accuracy=0.001)`, $\eta^2_G$ = `r round(RTAOVresults$ANOVA[2,7],digits=2)`) and stimming condition ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTAOVBF)$bf[1]),digits=2)`, *F* = `r round(RTAOVresults$ANOVA[1,4],digits=2)`, *p* = `r p_format(RTAOVresults$ANOVA[1,5],digits=1,accuracy=0.001)`, $\eta^2_G$ = `r round(RTAOVresults$ANOVA[1,7],digits=2)`), as well as an interaction between the two ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTAOVBF)$bf[3]),digits=2)`, *F* = `r round(RTAOVresults$ANOVA[3,4],digits=2)`, *p* = `r p_format(RTAOVresults$ANOVA[3,5],digits=1,accuracy=0.001)`, $\eta^2_G$ = `r round(RTAOVresults$ANOVA[3,7],digits=2)`). Note that reactions were faster in the non-stimming condition (`r meanRTnostimming` ms) than in the stimming condition (`r meanRTstimming` ms). These effects remained significant in the group who reported stimming (N=`r sum(reportsstimming)`) for the effects of cue congruency ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTAOVBFSTIM)$bf[2]),digits=2)`, *F* = `r round(RTAOVresultsSTIM$ANOVA[2,4],digits=2)`, *p* < `r p_format(RTAOVresultsSTIM$ANOVA[2,5],digits=1,accuracy=0.001)`, $\eta^2_G$ = `r round(RTAOVresultsSTIM$ANOVA[2,7],digits=2)`), stimming condition ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTAOVBFSTIM)$bf[1]),digits=2)`, *F* = `r round(RTAOVresultsSTIM$ANOVA[1,4],digits=2)`, *p* = `r p_format(RTAOVresultsSTIM$ANOVA[1,5],digits=2,accuracy=0.001)`, $\eta^2_G$ = `r round(RTAOVresultsSTIM$ANOVA[1,7],digits=2)`), and their interaction ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTAOVBFSTIM)$bf[3]),digits=2)`, *F* = `r round(RTAOVresultsSTIM$ANOVA[3,4],digits=2)`, *p* = `r p_format(RTAOVresultsSTIM$ANOVA[3,5],digits=2,accuracy=0.001)`, $\eta^2_G$ = `r round(RTAOVresultsSTIM$ANOVA[3,7],digits=2)`). For the group who did not report stimming (N=`r length(reportsstimming)-sum(reportsstimming)`), only the effect of cue congruency was significant ($log_{10}BF_{10}$ = `r round(log10(extractBF(RTAOVBFnoSTIM)$bf[2]),digits=2)`, *F* = `r round(RTAOVresultsNOSTIM$ANOVA[2,4],digits=2)`, *p* `r p_format(RTAOVresultsNOSTIM$ANOVA[2,5],digits=3,accuracy=0.001)`, $\eta^2_G$ = `r round(RTAOVresultsNOSTIM$ANOVA[2,7],digits=2)`).

```{r plotExpt2resps}
#| include: true
#| output: true
#| label: fig-Expt2resps
#| out.width: '95%'
#| fig.cap: 'Summary of reaction time and accuracy data for the cueing task in Experiment 2. Small points show data for individual participants (N = 22), and larger points give the group means. Error bars indicate 95% confidence intervals.'

knitr::include_graphics(paste0(figdir,'Exp2resps.pdf'))

```

```{r plotExpt2EEG}
#| include: true
#| output: true
#| label: fig-Expt2EEG
#| fig.cap: 'Summary of EEG data from Experiment 2. Panel (a) shows decoding accuracy of a pattern classification algorithm trained on responses to the cue stimulus (Time 0 is the cue onset) for the full sample of participants (N=22). Panel (b) shows decoding accuracy for classifying target location (black curve) or cue congruency (green curve) in response to the target (Time 0 is the target onset), again for all participants. Panels (c,d) compare decoding between the stimming (N=14) and non-stimming (N=8) participants for cue side and cue congruency. Panels (e,f) compare decoding accuracy for cue side within each group, and between the stimming and no stimming conditions. Shaded regions in all panels indicate 95% confidence intervals.'

knitr::include_graphics(paste0(figdir,'Exp2summary.pdf'))

```

# Discussion

# References